### 2024-12-17 doc/paper/paper.tex

\documentclass[12pt,a4paper]{article}
\input{preamble.tex}
\usepackage[margin=1in]{geometry}
\usepackage{authblk}
\externaldocument[s-]{supp}

\begin{document}
\title{Autopolyploid establishment through polygenic adaptation}
\author[1]{Arthur Zwaenepoel\thanks{arthur.zwaenepoel@uantwerpen.be}}
\affil[1]{\footnotesize Department of Biology, University of Antwerp, 2020 Antwerp, Belgium}
\date{\vspace{-5ex}}
\maketitle
\begin{abstract}
We define the infinitesimal model of quantitative genetics (\textit{sensu}
\cite{barton2017}) for the inheritance of an additive quantitative trait in a
mixed-ploidy population consisting of diploid, triploid and autotetraploid
individuals producing haploid and diploid gametes.
We implement efficient simulation methods and use these to study the
quantitative genetics of mixed-ploidy populations and the establishment of
autotetraploids in a new habitat.
We show that in the absence of migration from the source population,
autotetraploids are more likely to found a successful population than diploids
under a very broad range of conditions, but that this is unlikely to
sufficiently counter the scarcity of tetraploid founders when the source is
predominantly diploid.
[...]
\end{abstract}

\section*{Introduction}

% Should connect more with literature and concepts of isolating mechanisms pre
% vs. postzygotic isolation, and reinforcement etc. (see e.g. Parvathy &
% Himani's paper introduction)

Many plant species exhibit ploidy variation
\citep{levin2002,soltis2007,rice2015}, and many of these \textit{mixed-ploidy}
species have populations in which different cytotypes coexist or form contact
zones \citep{kolar2017}.
How such mixed-ploidy populations emerge and are maintained has proven somewhat
challenging to understand.

Consider for instance a randomly mating diploid population.
Under the commonly accepted view that polyploids mostly emerge through the
union of unreduced gametes \citep{bretagnolle1995,herben2016,kreiner2017b}, a
new tetraploid individual originating by a chance encounter of two unreduced
diploid gametes (an event occurring at an appreciable rate; \cite{kreiner2017})
is highly unlikely to establish a stable tetraploid subpopulation, as most of
its gametes will end up in unfit hybrids of odd ploidy level (triploid block,
\cite{ramsey1998,kohler2010,brown2024}).
This negative frequency dependence effect is commonly referred to as
\textit{minority cytotype exclusion} (MCE), after \cite{levin1975}.
It is well-appreciated that, due to MCE, in a large random mating mixed-ploidy
population dominated by diploids, the rate of unreduced gamete formation needs
to be extraordinarily high for tetraploids to establish (\cite{felber1997}, see
also \cref{s-sec:det}).

Hence, to explain the widespread occurrence of mixed-ploidy populations,
additional factors besides the continuous formation of polyploids through the
union of unreduced gametes need to be considered.
Firstly, chance establishment of tetraploids through drift could occur.
Indeed, the problem is somewhat analogous to the spread of underdominant
chromosomal rearrangements, where local establishment through genetic drift and
subsequent spreading in a subdivided population through local extinction and 
recolonization dynamics has been suggested as a plausible model
\citep{lande1985}.
However, MCE is quite strong in randomly mating mixed-ploidy populations, and
the population size has to be very small for local tetraploid establishment to
occur at an appreciable rate (\cite{rausch2005}, see also \cref{s-sec:mchain}).
Secondly, any form of prezygotic isolation between cytotypes could
promote establishment of polyploid cytotypes by alleviating MCE.
Particularly relevant are assortative mating by cytotype  (for instance through
phenological differences across cytotypes, or differences in pollinators;
\cite{kolar2017}), self-fertilization \citep{rausch2005,novikova2023}, and
localized dispersal \citep{baack2005,kolar2017}.
Finally, selection may be invoked to explain the establishment of polyploids.
Tetraploids may have higher relative fitness than their diploid counterparts
due to reduced inbreeding depression \citep{ronfort1999}, or due to being
better adapted to (changing) environmental conditions \citep{vandepeer2021}.

None of these factors is likely to explain by itself the establishment of
polyploids, and the consensus in the field appears to be that some mix of the
above is required to explain the occurrence of mixed-ploidy populations in
nature \citep{kolar2017,mortier2024}.
In particular, polyploids are thought to establish mainly in novel, unoccupied
habitats where they evade MCE (for instance at range edges, or after local
extinction due to environmental change).
However, such peripheral habitats are likely to present adaptive challenges to
establishment \citep{kawecki2008}, and if polyploids are able to colonize such
habitats at an appreciable rate, they must somehow be better adapted to local
conditions, or more able to adapt to those conditions despite inbreeding and
migration from the source population while the population is still small,
compared to diploids.
Indeed, when the source population is dominated by diploids, the probability
that a migrant individual is tetraploid will be very small ($O(u^2)$ if $u$ is
the probability that a diploid in the source population produces an unreduced
gamete in meiosis), so that tetraploids need a substantial advantage during
colonization if they are to establish before diploids do.

While there have been substantial modeling efforts aimed at understanding
autotetraploid establishment within diploid populations (e.g. \cite{levin1975,
felber1991, felber1997, rausch2005, oswald2011, clo2022c, griswold2021}), the
problem of polyploid establishment in a novel habitat which presents some adaptive
challenge remains largely unaddressed, despite its centrality to verbal
arguments about the establishment of polyploids in natural populations
\citep{kolar2017, vandepeer2021, clo2022d}.

Here we develop a model for the establishment of a mixed-ploidy population in a
novel, unoccupied habitat based on \cite{barton2018}.
In order to establish in the novel habitat, the population has to adapt to
local environmental conditions.
We assume fitness is determined by directional selection on a single polygenic
trait which can be interpreted as log fitness at low density in the new
habitat.
As in \cite{barton2018}, we assume the trait follows the infinitesimal
model (\textit{sensu} \cite{barton2017}, i.e. the `Gaussian descendants'
infinitesimal model \citep{turelli2017}).
We extend the infinitesimal model, and the approach for exact simulation of
trait evolution under the infinitesimal model, to mixed-ploidy populations.
We then use simulations to study tetraploid establishment, both from single
migrants and under continuous migration from a predominantly diploid source
population, examining the effects of autopolyploid genetics,
maladaptive migration, selfing and assortative mating on the probability that
autotetraploids establish in the novel habitat.

\section*{Model and Methods}

\begin{table}[t]
\caption{Glossary of the main notation used in the main text.
} \label{tbl:glossary}
\centering
\small
\begin{tabularx}{\linewidth}{lX}
\cline{1-2}
\textbf{notation}   & \textbf{description}   \\ \cline{1-2}
$N$ & total population size\\
$N_k$ & population size of the $k$-ploid cytotype \\
$\pi_k$ & equilibrium frequency of the $k$-ploid cytotype \\
$u$ & probability of unreduced gamete formation ($u=u_{22}=u_{32}=1-u_{42}$)\\
$v$ & probability that a triploid produces a haploid/diploid gamete
  ($v=u_{31}=u_{32}$)\\
$m$ & expected number of migrants per generation arriving in the new habitat \\
$z_i$ & trait value of individual $i$ \\
$c_i$ & ploidy level of individual $i$ \\
$g_i$ & ploidy level of gamete produced by individual $i$ in a particular cross\\
$V$ & segregation variance in the reference diploid population \\
$V_{i,k}$ & gametic segregation variance associated with the production of a
  $k$-ploid gamete by individual $i$ \\
$\Vx_k$ & genetic variance associated with a haploid genome in the $k$-ploid
  reference population (i.e. a $k$-ploid non-inbred population at HWLE) \\
$\beta_{k}$ & scaling factor for allelic effects in $k$-ploids \\
$F_i$ & inbreeding coefficient in individual $i$ \\
$\Phi_{ij}$ & coancestry coefficient for individuals $i$ and $j$ \\
$\alpha_k$ & probability that the two genes at a locus in a diploid gamete
  formed by a $k$-ploid individual descend from the same parental gene copy\\
$\gamma$ & strength of directional selection in the new habitat\\
$\theta$ & trait value beyond which the growth rate becomes positive in the new
habitat \\ $w_{ij}$ & fitness of parental pair $(i,j)$ \\
$w_{ij}^{kl}$ & expected fitness of offspring from parental pair $(i,j)$ when
$i$ contributes a $k$-ploid gamete and $j$ contributes a $l$-ploid gamete \\
$\sigma_k$ & selfing rate in $k$-ploids \\
$\rho_k$ & probability of assortative mating in $k$-ploids \\
\cline{1-2}
\end{tabularx}%
\end{table}

\subsection*{Mixed-ploidy population model}

We consider a mixed-ploidy population of size $N$ consisting of $N_2$ diploid,
$N_3$ triploid and $N_4=N-N_2-N_3$ tetraploid individuals.
We assume an individual of ploidy level $k$ forms haploid and diploid gametes
with proportions $u_{k1}$ and $u_{k2}$, as well as a proportion
$1-u_{k1}-u_{k2}$ inviable (e.g. aneuploid or polyploid) gametes.
The (relative) fecundity of a $k$-ploid individual is hence $u_{k1} + u_{k2}$.
Unless stated otherwise, we will assume 
\begin{equation}
    \begin{pmatrix} 
    u_{21} & u_{22} \\ 
    u_{31} & u_{32} \\ 
    u_{41} & u_{42} 
    \end{pmatrix} =
    \begin{pmatrix} 
    1-u & u \\
    v & v \\
    0 & 1-u
    \end{pmatrix} \label{eq:U}
\end{equation}
where $u$ is referred to as the proportion of unreduced gametes, and $2v$ is
the proportion of euploid gametes produced by a triploid individual.

When two individuals mate, we assume they produce gametes according to their
ploidy level (\cref{eq:U}), which randomly combine to produce offspring (which
may be inviable if one of the contributing gametes is inviable).
Intrinsic fitness disadvantages associated with particular zygotic ploidy
levels or cross types (e.g. modeling phenomena such as `triploid block') can be
straighforwardly included at this level.
An analysis of a deterministic model (i.e. where $N \rightarrow \infty$) for
the cytotype dynamics and equilibrium cytotype composition under random mating
is included in \cref{s-sec:det} (see also \cite{felber1997,kauai2024}).
The stochastic version for finite and constant $N$ is analyzed briefly in
\cref{s-sec:mchain}.

\subsection*{Infinitesimal model}

\paragraph*{The basic infinitesimal model.}

Consider a population which expresses a quantitative trait determined by a
large number of additive loci of small effect.
The infinitesimal model approximates the inheritance of such a trait by
assuming that the trait value $Z_{ij}$ of a random offspring from parents with
trait values $z_i$ and $z_j$ follows a Gaussian distribution with mean equal to
the midparent value and variance which is independent of the mean:
  \begin{align}
  Z_{ij} \sim \Normal\left(\frac{z_i + z_j}{2}, V_{ij}\right)
  \label{eq:inf}
  \end{align}
Here, $V_{ij}$ is referred to as the \textit{segregation variance} in family
$(i,j)$.
This is the variation generated among offspring from the same parental pair due
to random Mendelian segregation in meiosis.
This approximation can be justified as arising from the limit where the number
of loci determining the trait tends to infinity \citep{barton2017}.

An alternative, and for our purposes useful, way to characterize the model
is to write $Z_{ij} = Y_i + Y_j$, where $Y_i$ and $Y_j$
are independent Gaussian random variables $Y_i \sim \Normal\left(\frac{z_i}{2},
V_i\right)$ (and similarly for $Y_j$).
We refer to $Y_i$ as the (random) \textit{gametic value} of individual $i$, and
to $V_i$ as the \textit{gametic segregation variance} of individual $i$.
This formulation is helpful in that it highlights that Mendelian segregation
occurs independently in both parents when gametes are produced, which then
combine additively to determine the offspring trait value.
This model applies readily to an autopolyploid population expressing a trait
with infinitesimal genetics.
For instance, when an autotetraploid produces a diploid gamete, it will pass on
half its trait value to the gamete on average, with a variance determined by
the details of tetraploid meiosis (which are not, in general, identical to
those of diploid meiosis, see below).

Note that in a finite population, the segregation variance will decay over time
as the population becomes more inbred.
Indeed, Mendelian segregation generates less variation among the gametes
produced by individual $i$ when that individual is more inbred, as segregation
at homozygous loci does not generate any variation.
When $F_i$ is the inbreeding coefficient relative to some ancestral reference
population with gametic segregation variance $V$ (i.e. the probability that two
genes at a locus in individual $i$ sampled without replacement are identical by
descent), the gametic segregation variance of individual $i$ will be
$V_i = (1-F_i)V$.
This holds for both diploids and tetraploids (\cref{s-sec:tetinbred}, also
\cite{moody1993}).

\paragraph{Scaling of traits across ploidy levels.}

If we would somewhat naively assume that the allelic effects underlying an
additive quantitative trait are identical across ploidy levels, 
a tetraploid offspring from a cross between two diploids would yield, on average,
a trait value which is the sum of the parental trait values, with a
distribution that depends on the variance generated by the process producing
unreduced gametes.
This is not likely to reflect biological reality: tetraploids do not tend to
have, for instance, twice the size of their diploid progenitors on average.
Similarly, the genetic variance at Hardy-Weinberg and linkage equilibrium
(HWLE) in tetraploids will be twice that of their diploid counterparts under
such assumptions, which is similarly unrealistic \citep{clo2022}.

In order to account for this, we introduce a scaling factor $\beta_k$,
accounting for the effects of polyploidization \textit{per se} on trait
expression in $k$-ploids.
To introduce and interpret this parameter, we consider an $L$-locus additive
model, with two alleles (0 and 1) at each locus.
For a $k$-ploid individual, let $X_{i,j}$ be the allele at homolog $j$ of locus
$i$.
We assume the trait value is determined by
\begin{equation}
  z = \sum_{i=1}^L\sum_{j=1}^k a_{i,k} X_{i,j}
\end{equation}
Where $a_{i,k}$ is the allelic effect of the 1 allele at locus $i$ in
$k$-ploids.
The genetic variance at HWLE in $k$-ploids ($\tilde{V}_{z,k}$) will then be
\begin{equation}
  \tilde{V}_{z,k} = k\sum_{i=1}^L a_{i,k}^2 p_iq_i = k\Vx_{k}
\end{equation}
where we refer to $\Vx_{k}$ as the variance associated with a haploid genome in
$k$-ploids at HWLE.
Note that we also have $\tilde{V}_{z,2} = 2\Vx_{2} = 2V$ \citep{barton2017},
where $V$ is the segregation variance in the diploid population.
If we now assume $a_{i,k} = \beta_k a_{i,2}$, i.e. allelic effects in $k$-ploids
are as in diploids, but scaled by a factor $\beta_k$.
This implies that
\begin{equation}
  \frac{\tilde{V}_{z,k}}{\tilde{V}_{z,2}} 
  = \frac{k\Vx_{k}}{2\Vx_{2}} = \frac{k}{2} \beta_{k}^2
\end{equation}
and hence also that $\Vx_{k} = \beta_k^2 \Vx_{2} = \beta_k^2 V$.
These relations will also hold in the infinitesimal limit where $a_{i,2}
\rightarrow 0$ as $L\rightarrow \infty$.

\paragraph{Mixed-ploidy infinitesimal model.}

We can extend the infinitesimal model to the mixed-ploidy case, assuming that the
gametic value, on the diploid trait scale, associated with a $k$-ploid gamete ($k
\in \{1,2\}$) from individual $i$ of ploidy level $c_i \in \{2,3,4\}$ is a
Gaussian random variable $Y_{i,k}$ with distribution
  $$Y_{i,k} \sim \Normal\left(\frac{k}{c_i}\frac{z_i}{\beta_{c_i}}, V_{i,k}\right)$$
where $V_{i,k}$ is the gametic segregation variance associated with the
production of a $k$-ploid gamete by individual $i$.
The trait value of an individual originating from the union of a $k$-ploid
gamete of individual $i$ and an $l$-ploid gamete from individual $j$ is then
  $$Z_{ij}^{kl} = \beta_{k+l}\left(Y_{i,k} + Y_{j,l}\right)$$
i.e., $Z_{ij}^{kl}$ is a Gaussian random variate with distribution
\begin{equation}
  Z_{ij}^{kl} \sim \Normal\left(
    \beta_{k+l} \left(
          \frac{k}{c_i}\frac{z_i}{\beta_{c_i}} 
        + \frac{l}{c_j}\frac{z_j}{\beta_{c_j}}\right), 
    \beta_{k+l}^2 (V_{i,k} + V_{j,l})\right)
   \label{eq:oneline}
\end{equation}

For the production of diploid gametes by a $k$-ploid individual, the
segregation variance depends not only on the segregation variance in the base
population ($V$) and the inbreeding coefficient ($F$), but also on the detailed
assumptions on how these aberrant meiotic processes occur.
Importantly however, these details only affect the gametic segregation variance
through the quantity $\alpha_k$, which is the probability that a $k$-ploid
transmits two copies of the same homolog to a diploid gamete.
Note that $\alpha_4$, the probability that a diploid gamete of a tetraploid
individual carries two copies of the same homolog, is the probability of
\textit{double reduction} (e.g. \cite{lynch1998} p.57), and is upper bounded by
1/6 \citep{stift2008}.
The value of $\alpha_2$ depends on the relative frequency of unreduced gamete
formation through so-called \textit{first} and \textit{second division
restitution} \citep{bretagnolle1995,storme2013}.
We summarize the expressions for the gametic segregation variance in
\cref{tbl:segvar}.
Detailed derivations can be found in \cref{s-sec:segvar}.

\begin{table}[t]
\caption{Gametic segregation variance for haploid and diploid gametes produced
by the three cytotypes in the mixed-ploidy model. $F_i$ is the inbreeding
coefficient of individual $i$ (producing the gamete), whereas $\alpha_k$ is the
probability that a diploid gamete from a $k$-ploid individual receives two
copies of the same parental gene. Note that we have $\alpha_3 \le 1/4$ and
$\alpha_4 \le 1/6$.
} \label{tbl:segvar}
\centering
\small
\begin{tabularx}{\linewidth}{XXX}
\cline{1-3}
\textbf{cytotype}   & \textbf{haploid gamete variance} & \textbf{diploid gamete
variance}        \\ \cline{1-3} \\[-2.5ex]
diploid    & $\frac{1}{2}(1-F_i)V$    & $2\alpha_2(1-F_i)V$  \\ \\[-2.5ex]
triploid   & $\frac{2}{3}(1-F_i)V$     & $\frac{2}{3}(1 + 3\alpha_3)(1-F_i)V$ \\ \\[-2.5ex]
tetraploid & --                      & $(1+2\alpha_4)(1-F_i)V$             \\
\cline{1-3}
\end{tabularx}%
\end{table}

\paragraph{Recursions for inbreeding coefficients}

We can simulate the mixed-ploidy infinitesimal model for a finite population
through a straightforward extension of the approach outlined in
\cite{barton2017}, provided we can efficiently track inbreeding and coancestry
coefficients across the different ploidy levels.
Denoting the parents of individual $i$ by $k$ and $l$, the recursion for the
inbreeding coefficients in the mixed-ploidy case becomes
\begin{align}
    F_i &= \Phi_{kl} & \text{if } & c_i = 2 \nonumber \\ 
    F_i &= \frac{1}{3}\left(F_k^\ast + 2\Phi_{kl}\right) & \text{if } 
        & c_i = 3, g_k = 2, g_l = 1 \nonumber \\ 
    F_i &= \frac{1}{3}\left(F_l^\ast + 2\Phi_{kl}\right) & \text{if } 
        & c_i = 3, g_k = 1, g_l = 2 \nonumber \\ 
    F_i &= \frac1 6 (F_k^\ast + F_l^\ast + 4\Phi_{kl}) & \text{if } & c_i = 4
\end{align}
where $F_k^\ast = \alpha_{c_k} + (1-\alpha_{c_k})F_k$ (\cref{s-sec:tetinbred}).
The recursion for the coancestry coefficients in is given by
\begin{align}
    \Phi_{ii} &= \frac{1}{c_{i}} \left(1 + (c_i-1)F_i\right) \nonumber \\
    \Phi_{ij} &= \sum_k \sum_l P_{ik}P_{jl} \Phi_{kl} & i \ne j 
    \label{eq:coancestry}
\end{align}
where the sums are over individuals in the parental population, and where
$P_{ik} \in \{0, \frac1 3, \frac1 2, \frac2 3, 1\}$ is the probability that a
gene copy in $i$ is derived from parent $k$.

\subsection*{Establishment model}

Our model for the establishment of a population in an initially unoccupied
habitat is based on \cite{barton2018}.
We assume a large non-inbred `mainland' population at HWLE and cytotype
equilibrium, with $\Ex[z] = 0$ irrespective of the cytotype.
In generation $t$, $M(t)$ migrant individuals arrive on an island (the new
habitat) joining $N^\ast(t)$ resident individuals, where $M(t)$ is Poisson
distributed with mean $m$. 
The migrant individuals are assumed to be unrelated to the resident
individuals.
After migration in generation $t$, the $N(t) = N^\ast(t) + M(t)$ individuals
reproduce sexually, and the offspring thus produced survives until the next
generation with a probability determined by their trait value.
In the basic model, random selfing is allowed (but see below for a model with
self-incompatibility).
We assume the trait is under directional selection, with fitness is $w(z) =
e^{\gamma(z - \theta)}$, where $\gamma$ is the intensity of directional
selection and $\theta$ is the trait value for which the growth rate of the
population becomes positive.

Again following \cite{barton2018}, we simulate the model by first calculating
the fitness of each parental pair $(i,j)$, which is the expected fitness of
offspring of this pair
\begin{equation}
  w_{ij}
    = \sum_{k=1}^2\sum_{l=1}^2 w_{ij}^{kl}
    = \sum_{k=1}^2\sum_{l=1}^2 u_{c_{i},k}u_{c_{j},l}
        \Ex\left[e^{\gamma(Z_{ij}^{kl} - \theta)}\right]
\end{equation}
The expectation on the right hand side can be calculated from \cref{eq:oneline}
using the moment-generating function of the Gaussian.
Having calculated the $w_{ij}$, the number of offspring surviving into the next
generation is calculated as $N^\ast(t+1) = \sum_{i,j}w_{ij}/N(t)$.
Next, $N^\ast(t+1)$ offspring individuals are sampled by sampling parental
pairs and gametes proportional to $w_{ij}^{kl}$, and sampling a trait value
in accordance with \cref{eq:oneline}.

\subsection*{Self-fertilization and assortative mating}

We model partial self-fertilization by assuming that a proportion
$\sigma_{c_i}$ of the ovules of individual $i$ with ploidy level $c_i$ are
fertilized by self-pollen, while the remaining proportion $1-\sigma_{c_i}$ are
fertilized by randomly sampled pollen (which may be self-pollen with
probability $1/N$). 
That is, the expected number of offspring from individual $i$ as mother
surviving after selection is
\begin{equation}
\Ex[w_i] = \sigma_{c_i} w_{ii} +
  (1-\sigma_{c_i})\left[\frac{1}{N}\sum_{j=1}^N w_{ij}\right]
\end{equation}
We hence assume no pollen limitation (all outcrossing ovules are fertilized),
and no pollen discounting (the probability of being a father is unaffected by
an individual's selfing rate).
When modeling self-incompatibility, we assume there is no intrinsic
disadvantage to self-incompatibility, except when there is only a single
individual in the population, i.e.
\begin{equation}
\Ex[w_i] = \begin{cases}
    \frac{1}{N-1}\sum_{j \ne i }w_{ij} & \text{if } N > 1\\ 
    0 & \text{if } N=1 \end{cases}
\end{equation}

We model assortative mating by ploidy level in a similar way, assuming that a
fraction $\rho_{c_i}$ of the ovules of individual $i$ are fertilized by pollen
sampled from the $c_i$-ploid portion of the population, while a fraction
$1-\rho_{c_i}$ is fertilized by pollen randomly sampled from the entire
population.
\begin{equation}
\Ex[w_i] = \rho_{c_i} \frac{1}{N_{c_{i}}} \sum_{j=1}^N \delta_{c_i,c_j}w_{ij}
 + (1-\rho_{c_i})
\left[\frac{1}{N}\sum_{j=1}^N w_{ij}\right]
\end{equation}

\subsection*{Implementation and availability}

Individual-based simulations for the mixed-ploidy infinitesimal model were
implemented in Julia \citep{julia}.
Documented code and simulation notebooks are available at
\url{https://github.com/arzwa/InfGenetics}.

\section*{Results}

\subsection*{Autotetraploid and mixed-ploidy infinitesimal model}

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{/home/arthur_z/dev/InfGenetics/doc/img/tetfininf.pdf}
\caption{
The infinitesimal model in autotetraploids.
Comparisons are shown for the decay of the genetic variance ($V_z$) due to
inbreeding in exact simulations of the infinitesimal model in autotetraploids
against individual-based simulations of autotetraploid populations with $L$
unlinked additive loci determining the quantitative trait. 
(A) Simulations of a model without double reduction ($\alpha_4=0$).
(B) Simulations of a model with maximal double reduction ($\alpha_4=1/6$) (for
all loci in the finite $L$ simulations).
We show window-smoothed values for visual clarity, with observed genetic
variances averaged in windows of 20 generations every 10 generations.
The black line marks $e^{-t/4N}$. 
We assume $N=250$ and $V_z(0) = 1$.
\label{fig:vztet}}
\end{figure}

We evaluate the accuracy of the autotetraploid infinitesimal model as an
approximation to the evolution of a quantitative trait determined by $L$
additive loci.
We find that the infinitesimal model with inbreeding generally yields accurate
predictions for the evolution of the genetic variance when the number of loci
is sufficiently large ($L \ge 100$, say, \cref{fig:vztet,s-fig:fininf}).
Furthermore, we confirm that, in the absence of double reduction, the decay in
genetic variance due to inbreeding after a time $t$ is well-predicted by
$e^{-t/4N}$ (\cref{fig:vztet}A), as expected from the results of \cite{arnold2012}.
As predicted, double reduction leads to an immediate increase in genetic
variance, but leads to accelerated inbreeding, causing faster loss of genetic
variance in the long-term (\cref{fig:vztet,s-fig:fininf}).
As noted by \cite{arnold2012}, the loss of genetic variation in the presence of
double reduction cannot, however, be described by a single long-term effective
population size.

Simulations for the mixed-ploidy model further confirm the correctness of our
infinitesimal approximation and highlight the importance of assumptions
regarding the scaling of allelic effects across ploidy levels
($\beta$ parameters, \cref{s-fig:vz}).
Although inbreeding is slower in autotetraploids than in diploids for the same
population size, the tetraploid fraction of a diploid-dominated mixed-ploidy 
population will have an equal or higher average inbreeding coefficient
(\cref{s-fig:f1}).
This is because in such a population, triploid and tetraploid individuals
mostly arise from gametes formed by diploid individuals or polyploid
individuals with very recent diploid ancestry (about $1+u+2v$ generations ago
for tetraploids, and $1+\frac{2}{3}(u+2v)$ generations ago for triploids, see
\cref{s-sec:ttdip}), so that the polyploid subpopulations will show an average
relatedness similar to that of the diploid population and not evolve as an
isolated higher-ploidy population would.
A nonzero probability of producing IBD diploid gametes ($\alpha_k > 0$) will
then further increase the inbreeding coefficient in the tetraploid and triploid
fraction of the population relative to their diploid progenitors
(\cref{s-fig:f1}).
Therefore, as long as diploids dominate, the effect of harboring some fraction
of the gene pool in polyploid individuals on the rate of inbreeding is
negligible, and we find that the evolution of the inbreeding coefficient over
time is well predicted by $1-e^{-t/2N_e}$, where the inbreeding-effective
population size is, to first order in $u$, given by $(1-2u)N$
(\cref{s-sec:effsize}).

\subsection*{Establishment from a single individual}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{/home/arthur_z/dev/InfGenetics/doc/img/est1b.pdf}
\caption{
(A) Probability of establishment from a single diploid or tetraploid individual
with trait value $z=0$ for increasing selection intensity $\gamma$.
We assume $m=0$ and $u=0$, i.e. there is no migration, and no unreduced
gametes are produced. The trait is scaled in tetraploids so as to yield the
same genetic variance at HWLE ($\beta_4 = \sqrt{1/2}$)
(B) Probability of a tetraploid individual with trait value $z = 0$
successfully founding a population ($P_4$), relative to the probability for a
diploid individual with the same trait value ($P_2$).
The vertical dashed line marks $\beta_4=\sqrt{1/2}$, for which the variance at
HWLE is identical between diploids and autotetraploids.
(C) Probability of tetraploid establishment relative to the probability of
diploid establishment (on a $\log_{10}$ scale) across a range of values for
$\gamma$ and $\beta_4$. The vertical line again marks $\beta_4=\sqrt{1/2}$.
All results are estimated from 500.000 replicate simulations and assume
$\theta=2$.
\label{fig:est1}}
\end{figure}

Having established the validity of the mixed-ploidy infinitesimal model, we now
use it to study the establishment of polyploids in a marginal habitat to which
migrants from a mixed-ploidy source population are maladapted.

We first consider the establishment of a population from a single migrant
individual with trait value $z_0 = 0$.
We assume $u=0$ and compare the probability of establishment when the migrant
is diploid vs. tetraploid (\cref{fig:est1}).
As noted by \cite{barton2018}, the establishment probability (for a given
cytotype) depends essentially on two dimensionless parameters,
$\gamma\sqrt{2V}$ and $\theta/\sqrt{2V}$, corresponding to the intensity of
selection and the degree of maladaptation, respectively.
We shall scale our results accordingly, assuming $2V=1$ throughout, and shall
assume $\theta=2$, i.e. adaptation has to increase the trait value by two
standard deviations to achieve a positive growth rate.

We find that reduced inbreeding in tetraploids substantially increases the
establishment probability.
Indeed, in the case where allelic effects are scaled across ploidy levels
to yield the same equilibrium genetic variance ($\beta = \sqrt{1/2}$, so that
polyploid migrants are not more likely to have high fitness than diploid
migrants), the establishment probability for tetraploids can be almost five
times as high as in diploids depending on the selection gradient
(\cref{fig:est1}A).
As the segregation variance and initial trait value are the same across these
simulations, this is a consequence of the reduced rate of inbreeding, which
slows down the exhaustion of the genetic variance carried by the initial
migrant individual.

While the probability of establishment becomes lower as the strength of
selection becomes large, the establishment probability does not decrease
monotonically with $\gamma$, i.e. as selection becomes very weak, the
establishment probability also decreases.
This happens because, although the probability of surviving the first couple of
generations becomes higher when selection is weaker, adaptation (i.e. increase
in the trait mean) will be slower,
and the risk that genetic variation is exhausted due to inbreeding before the
population is able to reach a consistently positive growth rate is increased as
a result.

Evidently, the scaling of the genetic variance across ploidy levels has a
profound effect on the establishment probability, but only when $\beta$ is
close to 0.5 (i.e. individual alleles have almost half the effect size in
tetraploids compared to diploids) is the benefit of the slower rate of
inbreeding in tetraploids canceled (\cref{fig:est1}B,C).

\subsection*{Establishment under recurrent migration}

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{/home/arthur_z/dev/InfGenetics/doc/img/estwmigration1b.pdf}
\caption{
Establishment with recurrent migration.
(A) Expected time until a population is established in the novel habitat for
increasing rates of migration and different degrees of maladaptation
($\bar{z}_s$, the mean trait value in the source population). 
(B) Proportion of simulation replicates in which tetraploids established. 
The dashed horizontal line marks the proportion of tetraploid migrants (i.e.
the proportion of tetraploids at equilibrium in the source population).
The solid horizontal lines mark $\frac{p_4\pi_4}{(p_4\pi_4 + p_2\pi_2)}$, where $p_2$ and
$p_4$ are the establishment probabilities for a single diploid and a single
tetraploid individual in the absence of migration (as in \cref{fig:est1}, but
with unreduced gamete formation included).
All results are based on 10.000 replicate simulations.
We assume $\gamma=0.25, \theta=1.5$ and $u=v=0.05$.
\label{fig:estwmig}}
\end{figure}

We next consider establishment in the new habitat when there is a continuous
influx of migrants (at rate $m$) coming from a large, noninbred and
predominantly diploid source population at cytotype equilibrium.
In this setting, establishment is certain to happen eventually, and we are
interested in the probability that a tetraploid population establishes before a
diploid one does, as well as the effects of maladaptive gene flow and MCE
(which both increase with increasing migration rate) on the time until
establishment.

We hypothesized that two counteracting processes affect the probability of
autotetraploid establishment in this scenario.
On the one hand, increased migration will increase the probability that an
otherwise likely successful tetraploid migrant suffers from MCE in the early
generations while the population size is low, because migrants are likely to be
diploid.
On the other hand, tetraploids are strongly reproductively isolated from
a typical migrant, so that they are less prone to maladaptive swamping.
Hence, conditional on evading MCE, they should be able to adapt to the new
habitat at a rate which is not strongly affected by the migration rate.
This contrasts with diploids, which interbreed freely with maladapted
migrants, resulting in a pulling back of the trait mean towards that of the
source population.

As expected, we find that the time to establishment (of a population of either
ploidy level) first decreases with increasing migration as a result of a larger
influx of potentially succesful migrants, but later increases with increasing
migration as a result of swamping by gene flow (\cref{fig:estwmig}A).
The probability that a tetraploid population establishes decreases
monotonically with increasing migration and is considerably larger than the
probability that a migrant individual is tetraploid over a large part of the
parameter range (\cref{fig:estwmig}B).
Together, this suggests that the negative effects of MCE generally outweigh the
benefits that come from reproductive isolation as migration increases.

\subsection*{Loss of self-incompatibility and assortative mating}

When polyploidization disrupts an existing SI system, we expect that
tetraploids suffer less from MCE as some portion of their ovules are now
assured to be fertilized by diploid gametes.
At the same time, we expect that accelerated inbreeding in selfing tetraploids
diminishes the adaptive advantage of tetraploids that we found above.
As shown in \cref{fig:selfing}, we find that when polyploidization is
associated with the loss of a SI system, tetraploids have a strongly increases
establishment probability.
This is the case even when the selfing rate $\sigma$ in tetraploids is zero (in
which case, under our modeling assumptions, there is only random selfing, i.e.
the \textit{realized} selfing rate in tetraploids is $1/N$).
Besides reducing MCE, selfing is also expected to simultaneously reduce the
effects of maladaptive gene flow, and we indeed find that when the selfing rate
is sufficiently high ($\ge 0.4$ in \cref{fig:selfing}), the relative
establishment probability of tetraploids increases with increasing migration
rate.
In this regime, the increased strength of MCE due to increased migration is
compensated by the stronger maladaptive gene flow experienced by diploids.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{/home/arthur_z/dev/InfGenetics/doc/img/estwmigration-si2.pdf}
\caption{
Establishment with recurrent migration and loss of self-incompatibility in
polyploids (diploids are assumed to be self-incompatible).
(A) Expected time until a population is established in the novel habitat for
increasing rates of migration and different self-fertilization rates in
polyploids. Triploids and tetraploids have the same selfing rate. 
Diploids are assumed to be self-incompatible.  Note that $\sigma=0.0$ refers to
random self-fertilization (i.e.  self-fertilization with probability $1/N$). 
(B) Proportion of simulation replicates in which tetraploids established. 
All results are based on 25.000 replicate simulations.
We assume $\gamma=0.25, \theta=1.5$ and $u=v=0.05$.
\label{fig:selfing}}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{/home/arthur_z/dev/InfGenetics/doc/img/estwmigration-am2.pdf}
\caption{
Establishment with recurrent migration and assortative mating by cytotype.
(A) Expected time until a population is established in the novel habitat for
increasing rates of migration and different probabilities of assortative mating
($\rho_k = \rho$ for $k=2,3,4$, where $\rho_k$ is the probability that an ovule
from a $k$-ploid mother is pollinated by a $k$-ploid father).
(B) Proportion of simulation replicates in which tetraploids established. 
All results are based on 25.000 replicate simulations.
We assume $\gamma=0.25$ and $u=v=0.05$.
\label{fig:am}}
\end{figure}

Another prezygotic isolating mechanism that has often been considered relevant
for explaining tetraploid establishment is assortative mating by ploidy level,
where ovules from a tetraploid are more likely to be fertilized by
pollen coming from a tetraploid -- irrespective of the trait values of these
individuals.
Clearly, assortative mating increases the probability of tetraploid
establishment (\cref{fig:am}), although not as strongly as the loss of an SI
system does.
Again, we find that for some parameter values (roughly $\rho \ge 0.4$),
assortative mating may be strong enough so that tetraploid establishment
increases with increasing migration rates, suggesting that tetraploids evade
maladapive gene flow sufficiently to overcome MCE.

\section*{Discussion}

Despite a sizeable literature on the topic, the question whether the
establishment of autopolyploids requires particular fitness advantages
associated with polyploidy or not has remained rather unclear and a topic of
considerable speculation.
The observation that polyploid populations tend to inhabit more extreme
habitats or occur at range margins of their diploid progenitors has been
interpreted as supporting the hypothesis that polyploids may be more tolerant
to extreme environmental conditions.

In a recent study, \cite{griswold2021} also investigated the effects of
maladaptive gene flow on polyploid establishment in a peripheral habitat,
focusing however on the case where fitness is determined by a single biallelic
locus rather than a polygenic trait.
The author found that autotetraploids can establish when adaptation in the
peripheral habitat is conferred by recessive alleles, when inbreeding
depression in tetraploids is lower than in diploids, and when maladaptive gene
flow is due to pollen dispersal, rather than seed dispersal.

We have ignored the effects of dominance throughout. Polyploids are expected to
accumulate a larger mutation load when deleterious variation is recessive due
to less efficient purging, and this may yield increased inbreeding depression.
These effects have been studied in the context of range expansions
\citep{booker2024}.
However, this applies only to polyploids that have been established for a long
time.
In our case, polyploids are always recently descended from diploid ancestors,
so they will not have accumulated more deleterious recessives than their
diploid counterparts.
As inbreeding is slower in tetraploids, we should hence have reduced inbreeding
depression in tetraploids, since they have the same deleterious load as
diploids but it is less expressed (when recessive).  
There should therefore be an immediate benefit to polyploidy (reduced
inbreeding depression), which is however associated with a long term
disadvantage because eventually the established polyploid will start to
accumulate a higher load due to less efficient purging.

\citep{novikova2023}

\section*{Acknowledgements}

I thank Christelle Fraïsse for employment, encouragement and feedback. I
acknowledge funding from the Research Foundation -- Flanders (FWO, Junior
Postdoctoral Fellowship 1272625N).

\bibliographystyle{abbrvnat}
\bibliography{/home/arthur_z/vimwiki/bib.bib}
\end{document}