### 2025-01-14 doc/paper/paper.tex
paper update

\documentclass[12pt,a4paper]{article}
\input{preamble.tex}
\usepackage[margin=1in]{geometry}
\usepackage{authblk}
\externaldocument[s-]{supp}

\begin{document}
\title{Autopolyploid establishment through polygenic adaptation}
\author[1]{Arthur Zwaenepoel\thanks{arthur.zwaenepoel@uantwerpen.be}}
\affil[1]{\footnotesize Department of Biology, University of Antwerp, 2020 Antwerp, Belgium}
\date{\vspace{-5ex}}
\maketitle
\begin{abstract}
We define the infinitesimal model of quantitative genetics (\textit{sensu}
\cite{barton2017}) for the inheritance of an additive quantitative trait in a
mixed-ploidy population consisting of diploid, triploid and autotetraploid
individuals producing haploid and diploid gametes.
We implement efficient simulation methods and use these to study the
quantitative genetics of mixed-ploidy populations and the establishment of
autotetraploids in a new habitat.
We show that for rare migration from the source population, autotetraploids are
more likely to found a successful population than diploids under a very
broad range of conditions, but that this is unlikely to sufficiently
counter the scarcity of tetraploid founders when the source is
predominantly diploid.

\end{abstract}

\section*{Introduction}

% Should connect more with literature and concepts of isolating mechanisms pre
% vs. postzygotic isolation, and reinforcement etc. (see e.g. Parvathy &
% Himani's paper introduction)

Many plant species exhibit ploidy variation
\citep{levin2002,soltis2007,rice2015}, and many of these \textit{mixed-ploidy}
species have populations in which different cytotypes coexist or form contact
zones \citep{kolar2017}.
How such mixed-ploidy populations emerge and are maintained has proven somewhat
challenging to understand.

Consider for instance a randomly mating diploid population.
Under the commonly accepted view that new polyploid plants are mostly formed
through the union of unreduced gametes
\citep{bretagnolle1995,herben2016,kreiner2017b}, a new tetraploid individual
originating by a chance encounter of two unreduced diploid gametes (an event
occurring at an appreciable rate; \cite{kreiner2017}) is highly unlikely to
establish a stable tetraploid subpopulation, as most of its gametes will end up
in unfit hybrids of odd ploidy level (a phenomenon referred to as `triploid
block', see \cite{ramsey1998,kohler2010,brown2024}).
This negative frequency dependence effect is commonly referred to as
\textit{minority cytotype exclusion} (MCE), after \cite{levin1975}.
It is well-appreciated that, as a consequence of MCE, the rate of unreduced
gamete formation needs to be extraordinarily high for tetraploids to establish
in a large random mating initially diploid population (\cite{felber1997}, see
also \cref{s-sec:det}).

Hence, to explain the widespread occurrence of mixed-ploidy populations,
additional factors besides the continuous formation of polyploids through the
union of unreduced gametes need to be considered.
Firstly, chance establishment of tetraploids through drift could occur.
Indeed, the problem is somewhat analogous to the spread of underdominant
chromosomal rearrangements, where local establishment through genetic drift and
subsequent spreading in a subdivided population through local extinction and 
recolonization dynamics has been suggested as a plausible model
\citep{lande1985}.
However, MCE is quite strong in randomly mating mixed-ploidy populations, and
the population size has to be very small for local tetraploid establishment to
occur at an appreciable rate (\cite{rausch2005}, see also \cref{s-sec:mchain}).
Secondly, any form of \textit{prezygotic isolation} between cytotypes could
promote establishment of polyploid cytotypes by alleviating MCE.
Particularly relevant are assortative mating by cytotype  (for instance through
phenological differences across cytotypes, or differences in pollinators;
\cite{kolar2017}), self-fertilization \citep{rausch2005,novikova2023}, and
localized dispersal \citep{baack2005,kolar2017}.
Finally, selection may be invoked to explain the establishment of polyploids.
Tetraploids may have higher relative fitness than their diploid counterparts
due to reduced inbreeding depression \citep{ronfort1999}, or due to being
better adapted to (changing) environmental conditions \citep{vandepeer2021}.

None of these factors is likely to explain by itself the establishment of
polyploids, and the consensus in the field appears to be that some mix of the
above is required to explain the occurrence of mixed-ploidy populations in
nature \citep{kolar2017,mortier2024}.
In particular, polyploids are thought to establish mainly in novel, unoccupied
habitats where they evade MCE (for instance at range edges, or after local
extinction due to environmental change) \citep{griswold2021}.
However, such peripheral habitats are likely to present adaptive challenges to
establishment \citep{kawecki2008adaptation}, and if polyploids are able to
colonize such habitats at an appreciable rate, they must somehow be better
adapted to local conditions, or more able to adapt to those conditions despite
inbreeding and maladaptive migration, compared to diploids.
Indeed, when the source population is dominated by diploids, the probability
that a migrant individual is tetraploid will be very small ($O(u^2)$ if $u$ is
the probability that a diploid in the source population produces an unreduced
gamete in meiosis), so that tetraploids need a substantial advantage during
colonization if they are to establish before diploids do.

While there have been substantial modeling efforts aimed at understanding
autotetraploid establishment within diploid populations \citep{levin1975,
felber1991, felber1997, rausch2005, oswald2011, clo2022c}, the problem of
polyploid establishment in peripheral habitats remains largely unaddressed
(but see \cite{griswold2021}),
despite its centrality to verbal arguments about the establishment of
polyploids in natural populations \citep{kolar2017, vandepeer2021, clo2022d}.

Here we develop a rather general model for the establishment of a mixed-ploidy
population in a novel, unoccupied habitat based on \cite{barton2018}.
In order to establish in the novel habitat, the population has to adapt to
local environmental conditions.
We assume fitness is determined by directional selection on a single polygenic
trait which can be interpreted as log fitness at low density in the new
habitat.
As in \cite{barton2018}, we assume the trait follows the infinitesimal
model (\textit{sensu} \cite{barton2017}, i.e. the `Gaussian descendants'
infinitesimal model \citep{turelli2017}).
We extend the infinitesimal model, and the approach for exact simulation of
trait evolution under the infinitesimal model, to mixed-ploidy populations.
This is a first contribution of the present paper.
We then use simulations to study tetraploid establishment, both from single
migrants and under continuous migration from a predominantly diploid source
population, examining the effects of autopolyploid genetics,
maladaptive migration, selfing and assortative mating on the probability that
autotetraploids establish in the novel habitat.

\section*{Model and Methods}

\begin{table}[t]
\caption{Glossary of the notation used in the main text.
} \label{tbl:glossary}
\centering
\small
\begin{tabularx}{\linewidth}{lX}
\cline{1-2}
\textbf{notation}   & \textbf{description}   \\ \cline{1-2}
$N$ & total population size\\
$N_k$ & population size of the $k$-ploid cytotype \\
$\pi_k$ & equilibrium frequency of the $k$-ploid cytotype \\
$u$ & probability of unreduced gamete formation ($u=u_{22}=u_{32}=1-u_{42}$)\\
$v$ & probability that a triploid produces a haploid/diploid gamete
  ($v=u_{31}=u_{32}$)\\
$m$ & expected number of migrants per generation arriving in the new habitat \\
$z_i$ & trait value of individual $i$ \\
$c_i$ & ploidy level of individual $i$ \\
$g_i$ & ploidy level of gamete produced by individual $i$ in a particular cross\\
$V$ & segregation variance in the reference diploid population \\
$V_{i,k}$ & gametic segregation variance associated with the production of a
  $k$-ploid gamete by individual $i$ \\
$\Vx_k$ & genetic variance associated with a haploid genome in the $k$-ploid
  reference population (i.e. a $k$-ploid non-inbred population at HWLE) \\
$\beta_{k}$ & scaling factor for allelic effects in $k$-ploids \\
$F_i$ & inbreeding coefficient in individual $i$ \\
$\Phi_{ij}$ & coancestry coefficient for individuals $i$ and $j$ \\
$\alpha_k$ & probability that the two genes at a locus in a diploid gamete
  formed by a $k$-ploid individual descend from the same parental gene copy\\
$\gamma$ & strength of directional selection in the new habitat\\
$\theta$ & trait value beyond which the growth rate becomes positive in the new
habitat \\ $w_{ij}$ & fitness of parental pair $(i,j)$ \\
$w_{ij}^{kl}$ & expected fitness of offspring from parental pair $(i,j)$ when
$i$ contributes a $k$-ploid gamete and $j$ contributes a $l$-ploid gamete \\
$\sigma_k$ & rate of self-fertilization in $k$-ploids \\
$\rho_k$ & probability of assortative mating in $k$-ploids \\
\cline{1-2}
\end{tabularx}%
\end{table}

\subsection*{Mixed-ploidy population model}

We consider a mixed-ploidy population of size $N$ consisting of $N_2$ diploid,
$N_3$ triploid and $N_4=N-N_2-N_3$ tetraploid individuals.
We assume an individual of ploidy level $k$ forms haploid and diploid gametes
with proportions $u_{k1}$ and $u_{k2}$, as well as a proportion
$1-u_{k1}-u_{k2}$ inviable (e.g. aneuploid or polyploid) gametes.
The (relative) fecundity of a $k$-ploid individual is hence $u_{k1} + u_{k2}$.
Unless stated otherwise, we will assume 
\begin{equation}
    \begin{pmatrix} 
    u_{21} & u_{22} \\ 
    u_{31} & u_{32} \\ 
    u_{41} & u_{42} 
    \end{pmatrix} =
    \begin{pmatrix} 
    1-u & u \\
    v & v \\
    0 & 1-u
    \end{pmatrix} \label{eq:U}
\end{equation}
where $u$ is referred to as the proportion of unreduced gametes, and $2v$ is
the proportion of euploid gametes produced by a triploid individual.

When two individuals mate, we assume they produce gametes according to their
ploidy level (\cref{eq:U}), which randomly combine to produce offspring (which
may be inviable if one of the contributing gametes is inviable).
Intrinsic fitness disadvantages associated with particular zygotic ploidy
levels or cross types (e.g. modeling phenomena such as `triploid block') can be
straighforwardly included at this level.
An analysis of a deterministic model (i.e. where $N \rightarrow \infty$) for
the cytotype dynamics and equilibrium cytotype composition under random mating
is included in \cref{s-sec:det} (see also \cite{felber1997,kauai2024}).
The stochastic version for finite and constant $N$ is analyzed briefly in
\cref{s-sec:mchain}.

\subsection*{Infinitesimal model}

\paragraph*{The basic infinitesimal model.}

Consider a population which expresses a quantitative trait determined by a
large number of additive loci of small effect.
The infinitesimal model approximates the inheritance of such a trait by
assuming that the trait value $Z_{ij}$ of a random offspring from parents with
trait values $z_i$ and $z_j$ follows a Gaussian distribution with mean equal to
the midparent value and variance which is independent of the mean:
  \begin{align}
  Z_{ij} \sim \Normal\left(\frac{z_i + z_j}{2}, V_{ij}\right)
  \label{eq:inf}
  \end{align}
Here, $V_{ij}$ is referred to as the \textit{segregation variance} in family
$(i,j)$.
This is the variation generated among offspring from the same parental pair due
to random Mendelian segregation in meiosis.
This approximation can be justified as arising from the limit where the number
of loci determining the trait tends to infinity \citep{barton2017}.

An alternative, and for our purposes useful, way to characterize the model
is to write $Z_{ij} = Y_i + Y_j$, where $Y_i$ and $Y_j$
are independent Gaussian random variables $Y_i \sim \Normal\left(\frac{z_i}{2},
V_i\right)$ (and similarly for $Y_j$).
We refer to $Y_i$ as the (random) \textit{gametic value} of individual $i$, and
to $V_i$ as the \textit{gametic segregation variance} of individual $i$.
This formulation is helpful in that it highlights that Mendelian segregation
occurs independently in both parents when gametes are produced, which then
combine additively to determine the offspring trait value.
This model applies readily to an autopolyploid population expressing a trait
with infinitesimal genetics.
For instance, when an autotetraploid produces a diploid gamete, it will pass on
half its trait value to the gamete on average, with a variance determined by
the details of tetraploid meiosis (which are not, in general, identical to
those of diploid meiosis, see below).

In a finite population, the segregation variance will decay over time
as the population becomes more inbred.
Indeed, Mendelian segregation generates less variation among the gametes
produced by individual $i$ when that individual is more inbred, as segregation
at homozygous loci does not generate any variation.
When $F_i$ is the inbreeding coefficient relative to some ancestral reference
population with gametic segregation variance $V$ (i.e. the probability that two
genes at a locus in individual $i$ sampled without replacement are identical by
descent), the gametic segregation variance of individual $i$ will be
$V_i = (1-F_i)V$.
This holds for both diploids and tetraploids (\cref{s-sec:tetinbred}, also
\cite{moody1993}).

\paragraph{Scaling of traits across ploidy levels.}

If we would somewhat naively assume that the allelic effects underlying an
additive quantitative trait are identical across ploidy levels, 
a tetraploid offspring from a cross between two diploids would yield, on average,
a trait value which is the sum of the parental trait values, with a
distribution that depends on the variance generated by the process producing
unreduced gametes.
This is not likely to reflect biological reality: tetraploids do not tend to
have, for instance, twice the size of their diploid progenitors on average.
Similarly, the genetic variance at Hardy-Weinberg and linkage equilibrium
(HWLE) in a large non-inbred tetraploid population will be twice that of their
diploid counterparts under such assumptions, which is similarly unrealistic
\citep{clo2022}.

In order to account for this, we introduce a scaling factor $\beta_k$,
accounting for the effects of polyploidization \textit{per se} on trait
expression in $k$-ploids.
To introduce and interpret this parameter, we consider an $L$-locus additive
model, with two alleles (0 and 1) at each locus.
For a $k$-ploid individual, let $X_{i,j}$ be the allele at homolog $j$ of locus
$i$.
We assume the trait value is determined by
\begin{equation}
  z = \sum_{i=1}^L\sum_{j=1}^k a_{i,k} X_{i,j}
\end{equation}
Where $a_{i,k}$ is the allelic effect of the 1 allele at locus $i$ in
$k$-ploids.
The genetic variance at HWLE in $k$-ploids ($\tilde{V}_{z,k}$) will then be
\begin{equation}
  \tilde{V}_{z,k} = k\sum_{i=1}^L a_{i,k}^2 p_iq_i = k\Vx_{k}
\end{equation}
where we refer to $\Vx_{k}$ as the variance associated with a haploid genome in
$k$-ploids at HWLE.
If we now assume $a_{i,k} = \beta_k a_{i,2}$, i.e. allelic effects in $k$-ploids
are as in diploids, but scaled by a factor $\beta_k$, and assume equal allele
frequencies in the different cytotypes, 
we find that
\begin{equation}
  \frac{\tilde{V}_{z,k}}{\tilde{V}_{z,2}} 
  = \frac{k\Vx_{k}}{2\Vx_{2}} = \frac{k}{2} \beta_{k}^2
\end{equation}
Note that by definition $\beta_2=1$.
Under the infinitesimal model (where $a_{i,2} \rightarrow 0$ as $L\rightarrow
\infty$), we have $\tilde{V}_{z,2} = 2\Vx_{2} = 2V$ \citep{barton2017}, where
$V$ is the segregation variance in the diploid population.
Hence, in the infinitesimal limit we have $\Vx_{k} = \beta_k^2 \Vx_{2} =
\beta_k^2 V$.

\paragraph{Mixed-ploidy infinitesimal model.}

We can extend the infinitesimal model to the mixed-ploidy case, assuming that the
gametic value, on the diploid trait scale, associated with a $k$-ploid gamete ($k
\in \{1,2\}$) from individual $i$ of ploidy level $c_i \in \{2,3,4\}$ is a
Gaussian random variable $Y_{i,k}$ with distribution
\begin{equation}
    Y_{i,k} \sim \Normal\left(\frac{k}{c_i}\frac{z_i}{\beta_{c_i}}, V_{i,k}\right)
    \label{eq:gamete}
\end{equation}
where $V_{i,k}$ is the gametic segregation variance associated with the
production of a $k$-ploid gamete by individual $i$ (see below).
The trait value of an individual originating from the union of a $k$-ploid
gamete of individual $i$ and an $l$-ploid gamete from individual $j$ is then
  $$Z_{ij}^{kl} = \beta_{k+l}\left(Y_{i,k} + Y_{j,l}\right)$$
i.e., $Z_{ij}^{kl}$ is a Gaussian random variate with distribution
\begin{equation}
  Z_{ij}^{kl} \sim \Normal\left(
    \beta_{k+l} \left(
          \frac{k}{c_i}\frac{z_i}{\beta_{c_i}} 
        + \frac{l}{c_j}\frac{z_j}{\beta_{c_j}}\right), 
    \beta_{k+l}^2 (V_{i,k} + V_{j,l})\right)
   \label{eq:oneline}
\end{equation}

For the production of diploid gametes by a $k$-ploid individual, the
segregation variance depends not only on the segregation variance in the base
population ($V$) and the inbreeding coefficient ($F$), but also on the detailed
assumptions on how these aberrant meiotic processes occur.
Importantly however, these details only affect the gametic segregation variance
through the quantity $\alpha_k$, which is the probability that a $k$-ploid
transmits two copies of the same homolog to a diploid gamete.
Note that $\alpha_4$, the probability that a diploid gamete of a tetraploid
individual carries two copies of the same homolog, is the probability of
\textit{double reduction} (e.g. \cite{lynch1998} p.57), and is upper bounded by
1/6 \citep{stift2008}.
The value of $\alpha_2$ depends on the relative frequency of unreduced gamete
formation through so-called \textit{first} and \textit{second division
restitution} \citep{bretagnolle1995,storme2013}.
We summarize the expressions for the gametic segregation variance in
\cref{tbl:segvar}.
Detailed derivations can be found in \cref{s-sec:segvar}.

\begin{table}[t]
\caption{Gametic segregation variance for haploid and diploid gametes produced
by the three cytotypes in the mixed-ploidy model. $F_i$ is the inbreeding
coefficient of individual $i$ (producing the gamete), whereas $\alpha_k$ is the
probability that a diploid gamete from a $k$-ploid individual receives two
copies of the same parental gene. We assume $\alpha_3 \le 1/4$ and
    $\alpha_4 \le 1/6$ (see \cref{s-sec:segvar}).
} \label{tbl:segvar}
\centering
\small
\begin{tabularx}{\linewidth}{XXX}
\cline{1-3}
\textbf{cytotype}   & \textbf{haploid gamete variance} & \textbf{diploid gamete
variance}        \\ \cline{1-3} \\[-2.5ex]
diploid    & $\frac{1}{2}(1-F_i)V$    & $2\alpha_2(1-F_i)V$  \\ \\[-2.5ex]
triploid   & $\frac{2}{3}(1-F_i)V$     & $\frac{2}{3}(1 + 3\alpha_3)(1-F_i)V$ \\ \\[-2.5ex]
tetraploid & --                      & $(1+2\alpha_4)(1-F_i)V$             \\
\cline{1-3}
\end{tabularx}%
\end{table}

Writing $\bar{z}_2$ for the mean trait value in the diploid subpopulation,
\cref{eq:oneline} implies that a tetraploid offspring from a random diploid
parental pair has an expected trait value equal to $\bar{z}_4 =
2\beta_4\bar{z}_2$.
This hence implies that $|\bar{z}_4| \ge |\bar{z}_2|$, with equality only when
$\beta_4=1/2$ or $\bar{z}_2 = \bar{z}_4 = 0$. 
In other words, when $\beta_4 > 1/2$, we would have for all but $\bar{z}_2 = 0$
that newly formed tetraploids have more extreme phenotypes on average than
their diploid parents.
In our establishment model (see below), we shall therefore always consider 
the case where $\bar{z}_2=0$ in the source population, and think of the trait
value modeled as the deviation from the mean phenotype in the source
population.

The property that polyploid phenotypes are more extreme on average makes sense
if we consider the underlying genetic model: if we assume the source population
is at an equilibrium between mutation, drift and stabilizing selection,
segregating genetic variants that affect the trait will be equally likely to
have positive or negative allelic effects \citep{hayward2022}.
At equilibrium, any diploid individual with trait value $z - \bar{z}_2 > 0$
then carries an excess of variants with positive allelic effects on the trait,
and this excess should be exaggerated in its polyploid offspring, which carry
\textit{twice} the excess of their parent on average, unless allelic effects are
scaled exactly by one half (i.e. $\beta_4 = 1/2$). 
In other words, the mixed-ploidy model is not `coordinate free' as the basic
infinitesimal model is: \cref{eq:oneline} only makes sense when the trait
values that are modeled correspond to deviations from the mean values
associated with an underlying equilibrium state.

\paragraph{Recursions for inbreeding coefficients}

We can simulate the mixed-ploidy infinitesimal model for a finite population
through a straightforward extension of the approach outlined in
\cite{barton2017}, provided we can efficiently track inbreeding and coancestry
coefficients across the different ploidy levels.
Denoting the parents of individual $i$ by $k$ and $l$, the recursion for the
inbreeding coefficients in the mixed-ploidy case becomes
\begin{align}
    F_i &= \Phi_{kl} & \text{if } & c_i = 2 \nonumber \\ 
    F_i &= \frac{1}{3}\left(F_k^\ast + 2\Phi_{kl}\right) & \text{if } 
        & c_i = 3, g_k = 2, g_l = 1 \nonumber \\ 
    F_i &= \frac{1}{3}\left(F_l^\ast + 2\Phi_{kl}\right) & \text{if } 
        & c_i = 3, g_k = 1, g_l = 2 \nonumber \\ 
    F_i &= \frac1 6 (F_k^\ast + F_l^\ast + 4\Phi_{kl}) & \text{if } & c_i = 4
\end{align}
where $F_k^\ast = \alpha_{c_k} + (1-\alpha_{c_k})F_k$ (\cref{s-sec:tetinbred}).
The recursion for the coancestry coefficients in is given by
\begin{align}
    \Phi_{ii} &= \frac{1}{c_{i}} \left(1 + (c_i-1)F_i\right) \nonumber \\
    \Phi_{ij} &= \sum_k \sum_l P_{ik}P_{jl} \Phi_{kl} & i \ne j 
    \label{eq:coancestry}
\end{align}
where the sums are over individuals in the parental population, and where
$P_{ik} \in \{0, \frac1 3, \frac1 2, \frac2 3, 1\}$ is the probability that a
gene copy in $i$ is derived from parent $k$.

\subsection*{Establishment model}

Our model for the establishment of a population in an initially unoccupied
habitat is based on \cite{barton2018}.
We assume a large non-inbred `mainland' mixed-ploidy population at HWLE and
cytotype equilibrium, with $\Ex[z] = 0$ irrespective of the cytotype.
The equilibrium trait value distribution for the different cytotypes on the
mainland is complicated in general, but a very accurate approximation (which we
use throughout) is readily obtained for the case where $u$ is small (see
\cref{s-sec:mixeq}).

In generation $t$, $M(t)$ migrant individuals arrive on an island (the new
habitat) joining $N^\ast(t)$ resident individuals, where $M(t)$ is Poisson
distributed with mean $m$. 
The migrant individuals are assumed to be unrelated to the resident
individuals.
After migration in generation $t$, the $N(t) = N^\ast(t) + M(t)$ individuals
reproduce sexually, and the offspring thus produced survives until the next
generation with a probability determined by their trait value.
In the basic model, random selfing is allowed (but see below for a model with
self-incompatibility).
We assume the trait is under directional selection, with fitness $w(z) =
e^{\gamma(z - \theta)}$, where $\gamma$ is the intensity of directional
selection and $\theta$ is the trait value for which the growth rate of the
population becomes positive.
Throughout the paper, we define establishment as reaching a population of size
100.

Again following \cite{barton2018}, we simulate the model by first calculating
the fitness of each parental pair $(i,j)$, which is the expected fitness of
offspring of this pair
\begin{equation}
  w_{ij}
    = \sum_{k=1}^2\sum_{l=1}^2 w_{ij}^{kl}
    = \sum_{k=1}^2\sum_{l=1}^2 u_{c_{i},k}u_{c_{j},l}
        \Ex\left[e^{\gamma(Z_{ij}^{kl} - \theta)}\right]
\end{equation}
The expectation on the right hand side can be calculated from \cref{eq:oneline}
using the moment-generating function of the Gaussian.
Having calculated the $w_{ij}$, the number of offspring surviving into the next
generation is calculated as $N^\ast(t+1) = \sum_{i,j}w_{ij}/N(t)$.
Next, $N^\ast(t+1)$ offspring individuals are sampled by sampling parental
pairs and gametes proportional to $w_{ij}^{kl}$, and sampling a trait value
in accordance with \cref{eq:oneline}.

\subsection*{Self-fertilization and assortative mating}

We model partial self-fertilization by assuming that a proportion
$\sigma_{c_i}$ of the ovules of individual $i$ with ploidy level $c_i$ are
fertilized by self-pollen, while the remaining proportion $1-\sigma_{c_i}$ are
fertilized by randomly sampled pollen (which may be self-pollen with
probability $1/N$). 
That is, the expected number of offspring from individual $i$ as mother
surviving after selection is
\begin{equation}
\Ex[w_i] = \sigma_{c_i} w_{ii} +
  (1-\sigma_{c_i})\left[\frac{1}{N}\sum_{j=1}^N w_{ij}\right]
\end{equation}
We hence assume no pollen limitation (all outcrossing ovules are fertilized),
and no pollen discounting (the probability of being a father is unaffected by
an individual's selfing rate).
When modeling self-incompatibility, we assume there is no intrinsic
disadvantage to self-incompatibility, except when there is only a single
individual in the population, i.e.
\begin{equation}
\Ex[w_i] = \begin{cases}
    \frac{1}{N-1}\sum_{j \ne i }w_{ij} & \text{if } N > 1\\ 
    0 & \text{if } N=1 \end{cases}
\end{equation}

We model assortative mating by ploidy level in a similar way, assuming that a
fraction $\rho_{c_i}$ of the ovules of individual $i$ are fertilized by pollen
sampled from the $c_i$-ploid portion of the population, while a fraction
$1-\rho_{c_i}$ is fertilized by pollen randomly sampled from the entire
population.
\begin{equation}
\Ex[w_i] = \rho_{c_i} \frac{1}{N_{c_{i}}} \sum_{j=1}^N \delta_{c_i,c_j}w_{ij}
 + (1-\rho_{c_i})
\left[\frac{1}{N}\sum_{j=1}^N w_{ij}\right]
\end{equation}

\subsection*{Implementation and availability}

Individual-based simulations for the mixed-ploidy infinitesimal model were
implemented in Julia \citep{julia}.
Documented code and simulation notebooks are available at
\url{https://github.com/arzwa/InfGenetics}.

\section*{Results}

The results section is organized as follows: first we verify the correctness
and accuracy of the infinitesimal model for mixed-ploidy populations by
comparison against simulations of a model with finitely many loci.
Next we assess how autotetraploid genetics and different assumptions on the
genetic variance of neotetraploids affect the probability of tetraploid
establishment in a marginal habitat relative to diploid establishment, starting
from a single maladapted migrant individual.
We then study how continuous migration from a predominantly diploid maladapted
source population affects the relative establishment probability, assessing the
impact of migration load and MCE.
Finally, we consider the impact of prezygotic isolation mechanisms on
tetraploid establishment in the marginal habitat.
Throughout, we assume $\beta_3 = \sqrt{2/3}, \beta_4=\sqrt{1/2}$
and $\alpha_2 = \alpha_3 = \alpha_4 = 0$, unless stated otherwise.

\subsection*{Autotetraploid and mixed-ploidy infinitesimal model}

\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{/home/arzwa/dev/InfGenetics/doc/img/tetfininf.pdf}
\caption{
The infinitesimal model in autotetraploids.
Comparisons are shown for the decay of the genetic variance ($V_z$) due to
inbreeding in exact simulations of the infinitesimal model in autotetraploids
against individual-based simulations of autotetraploid populations with $L$
unlinked additive loci determining the quantitative trait. 
(A) Simulations of a model without double reduction ($\alpha_4=0$).
(B) Simulations of a model with maximal double reduction ($\alpha_4=1/6$) (for
all loci in the finite $L$ simulations).
We show window-smoothed values for visual clarity, with observed genetic
variances averaged in windows of 20 generations every 10 generations.
The black line marks $e^{-t/4N}$. 
We assume $N=250$ and $V_z(0) = 1$.
\label{fig:vztet}}
\end{figure}

We evaluate the accuracy of the autotetraploid infinitesimal model as an
approximation to the evolution of a quantitative trait determined by $L$
additive loci.
We find that the infinitesimal model with inbreeding generally yields accurate
predictions for the evolution of the genetic variance when the number of loci
is sufficiently large ($L \ge 100$, say, \cref{fig:vztet,s-fig:fininf}).
Furthermore, we confirm that, in the absence of double reduction, the decay in
genetic variance due to inbreeding after a time $t$ is well-predicted by
$e^{-t/4N}$ (\cref{fig:vztet}A), as expected from the results of \cite{arnold2012}.
As predicted, double reduction leads to an immediate increase in genetic
variance (as it increases the segregation variance), but leads to accelerated
inbreeding, causing faster loss of genetic variance in the long-term
(\cref{fig:vztet,s-fig:fininf}).
As noted by \cite{arnold2012}, the loss of genetic variation in the presence of
double reduction cannot, however, be described by a single long-term effective
population size.

Simulations for the mixed-ploidy model further confirm the correctness of our
infinitesimal approximation (\cref{s-fig:vz}), yielding accurate predictions
for the evolution of the phenotypic variance across ploidy levels and for
different assumptions regarding the scaling of allelic effects ($\beta$
parameters) and unreduced gamete formation ($\alpha$ parameters).

It is worth noting that, although inbreeding is slower in autotetraploids than
in diploids for the same population size, the tetraploid fraction of a
diploid-dominated mixed-ploidy population will have an equal or higher average
inbreeding coefficient (\cref{s-fig:f1}).
This is because in such a population, triploid and tetraploid individuals
mostly arise from gametes formed by diploid individuals, or by polyploid
individuals with very recent diploid ancestry (on average $1+u+2v$ generations ago
for tetraploids, and $1+\frac{2}{3}(u+2v)$ generations ago for triploids, see
\cref{s-sec:ttdip}).
As a result, the polyploid subpopulations will show an average relatedness
similar to that of the diploid population and not evolve as an isolated
higher-ploidy population would.
A nonzero probability of producing IBD diploid gametes ($\alpha_k > 0$) will
then further increase the inbreeding coefficient in the tetraploid and triploid
fraction of the population relative to their diploid progenitors
(\cref{s-fig:f1}).
Therefore, as long as diploids dominate, harboring some fraction of the gene
pool in polyploid individuals has a negligible effect on the rate of inbreeding
in the mixed-ploidy population as a whole, and we find that the evolution of
the inbreeding coefficient over time is well predicted by $1-e^{-t/2N_e}$,
where the inbreeding-effective population size is, to first order in $u$, given
by $(1-2u)N$ (\cref{s-sec:effsize}). This is just the expected number of
diploid individuals (to first order in $u$), highlighting that when diploids
dominate, polyploids have low fitness and do not contribute to the effective
population size.

\subsection*{Establishment from a single individual}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{/home/arzwa/dev/InfGenetics/doc/img/est1b.pdf}
\caption{
(A) Probability of establishment from a single diploid or tetraploid individual
with trait value $z=0$ for increasing selection intensity $\gamma$.
We assume $m=0$ and $u=0$, i.e. there is no migration, and no unreduced
gametes are produced. The trait is scaled in tetraploids so as to yield the
same genetic variance at HWLE ($\beta_4 = \sqrt{1/2}$)
(B) Probability of a tetraploid individual with trait value $z = 0$
successfully founding a population ($P_4$), relative to the probability for a
diploid individual with the same trait value ($P_2$).
The vertical dashed line marks $\beta_4=\sqrt{1/2}$, for which the variance at
HWLE is identical between diploids and autotetraploids.
(C) Probability of tetraploid establishment relative to the probability of
diploid establishment (on a $\log_{10}$ scale) across a range of values for
$\gamma$ and $\beta_4$. The vertical line again marks $\beta_4=\sqrt{1/2}$.
All results are estimated from 500.000 replicate simulations and assume
$\theta=2$. Establishment is defined as reaching $N=100$.
\label{fig:est1}}
\end{figure}

Having established the validity of the mixed-ploidy infinitesimal model, we now
use it to study the establishment of polyploids in a marginal habitat to which
migrants from a mixed-ploidy source population are maladapted.

We first consider the establishment of a population from a single migrant
individual with trait value $z_0 = 0$.
We assume $u=0$ (i.e. there are no unreduced gametes, and hence no newly formed
polyploids) and compare the probability of establishment when the migrant
is diploid vs. tetraploid (\cref{fig:est1}).
As noted by \cite{barton2018}, the establishment probability (for a given
cytotype) depends essentially on two dimensionless parameters,
$\gamma\sqrt{2V}$ and $\theta/\sqrt{2V}$, corresponding to the intensity of
selection and the degree of maladaptation, respectively.
We shall scale our results accordingly, assuming $2V=1$ throughout, and shall
assume $\theta=2$, so that the average trait value has to increase by two
standard deviations to achieve a positive growth rate.
The \textit{relative} establishment probability for different cytotypes will of
course also depend on how allelic effects scale across ploidy levels ($\beta$).

We find that reduced inbreeding in tetraploids substantially increases the
establishment probability of tetraploids relative to diploids across a large
part of the parameter space (\cref{fig:est1}).
Indeed, in the case where allelic effects are scaled across ploidy levels
to yield the same equilibrium genetic variance ($\beta_4 = \sqrt{1/2}$, so that
polyploid migrants are not more likely to have high fitness than diploid
migrants), the establishment probability for tetraploids can be almost five
times as high as in diploids depending on the selection gradient
(\cref{fig:est1}A).
As the segregation variance and initial trait value are the same across these
simulations, this is a consequence only of the reduced rate of inbreeding,
which slows down the exhaustion of the genetic variance carried by the initial
migrant individual.

While the probability of establishment becomes lower as the strength of
selection becomes large, the establishment probability does not decrease
monotonically with $\gamma$, i.e. as selection becomes very weak, the
establishment probability also decreases.
This happens because, although the probability of surviving the first couple of
generations becomes higher when selection is weaker, adaptation (i.e. increase
in the trait mean) will be slower,
increasing the risk that genetic variation is exhausted due to inbreeding before the
population is able to reach a consistently positive growth rate.
Note that when $\gamma = 0$, each individual has fitness 1 and we have a
critical branching process with a Poisson offspring distribution, so that the
probability to reach $N=100$ is $1/100$ \citep{barton2018}.

Evidently, the scaling of the genetic variance across ploidy levels has a
profound effect on the establishment probability, but only when $\beta_4$ is
close to 0.5 (i.e. individual alleles have almost half the effect size in
tetraploids compared to diploids) is the benefit of the slower rate of
inbreeding in tetraploids canceled (\cref{fig:est1}B,C).

\subsection*{Establishment under recurrent migration}

\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{/home/arzwa/dev/InfGenetics/doc/img/estwmigration-dr.pdf}
\caption{
    Establishment with recurrent migration.
    (A) Expected time until a population is established in the novel habitat
    for increasing rates of migration and different degrees of maladaptation
    ($\bar{z}_s$, the mean trait value in the source population). Results are
    shown for the case with $\alpha_k = 0$ for $k=2,3,4$. 
    (B) Proportion of simulation replicates in which tetraploids established.
    The dots connected by solid lines show simulation results with
    $\alpha_k=0$, whereas the dots connected by dashed lines
    show simulation results with $\alpha_2=1/2, \alpha_3=1/4$ and
    $\alpha_4=1/6$ (i.e. maximum $\alpha$).
    The solid horizontal lines mark $\frac{p_4\pi_4}{(p_4\pi_4 + p_2\pi_2)}$,
    where $p_2$ and $p_4$ are the establishment probabilities for a single
    diploid and a single tetraploid individual in the absence of migration (as
    in \cref{fig:est1}, but with unreduced gamete formation included). 
    The dashed horizontal line marks the proportion of tetraploid migrants
    (i.e. the proportion of tetraploids at equilibrium in the source
    population, $\approx 0.3$\%).
    The baseline predictions (horizontal lines) are based on $5\times 10^6$
    simulation replicates.
    All other results are based on 50.000 replicate simulations.
    We assume $\gamma=0.25$ and $u=v=0.05$.
    \label{fig:estwmig}}
\end{figure}

We next consider establishment in the new habitat when there is a continuous
influx of migrants ($m$ migrants per generation on average) coming from a large,
noninbred and predominantly diploid source population at cytotype equilibrium.
In this setting, establishment is certain to happen eventually, and we are
interested in the probability that a tetraploid population establishes before a
diploid one does.
More specifically, we are interested in how maladaptive gene flow and MCE
(which both increase with $m$) affect the relative establishment probability
and the time until establishment.

We hypothesized that two counteracting processes affect the probability of
autotetraploid establishment in this scenario.
On the one hand, increased migration will increase the probability that an
otherwise likely successful tetraploid migrant suffers from MCE in the early
generations while the population size is low, because migrants are likely to be
diploid.
On the other hand, tetraploids are more strongly reproductively isolated from
a typical migrant, so that a tetraploid subpopulation should be less prone to
maladaptive gene flow.
Hence, conditional on evading MCE, they should be able to adapt to the new
habitat at a rate which is not strongly affected by the migration rate.
This contrasts with diploids, which interbreed freely with maladapted
migrants, resulting in a pulling back of the trait mean towards that of the
source population.
However, MCE is expected to be stronger than the latter effect,
precluding the establishment of tetraploids as $m$ increases.
Lastly, as the mean trait value on the island increases in diploids
during adaptation, newly formed tetraploids on the island will, on average,
have more extreme phenotypes than diploid offspring when $\beta_4 > 1/2$, which
may also aid their establishment (irrespective of $m$).

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{/home/arzwa/dev/InfGenetics/doc/img/estwmigration1d.pdf}
\caption{
    Evolution of the mean inbreeding coefficient and trait value across
    establishment replicates, averaged by population size, for increasing rates
    of migration ($m$) from the predominantly diploid source population. 
    All results are based on 1000 succesful establishment replicates.
    We assume equal equilibrium variance across ploidy levels and $\gamma=0.25,
    \theta=2, 2V=1$ and $u=v=0.05$.
    The $m=0$ simulations are seeded with a single diploid or tetraploid
    individual with trait value 0, and assume $u=v=0$.
\label{fig:estwmig2}}
\end{figure}

As expected, we find that the time to establishment (of a population of either
ploidy level) first decreases with increasing migration as a result of a larger
influx of potentially succesful migrants, but later increases with increasing
migration as a result of swamping by gene flow (\cref{fig:estwmig}A).
Importantly, the tetraploid establishment probability is considerably larger
than the expected proportion of tetraploid migrants over a large part of the
parameter range (\cref{fig:estwmig}B, black dashed line).
Furthermore, we find that for small migration rates, the probability that
tetraploids establish rather than diploids can be larger than expected based on
the establishment probabilities in the absence of migration (colored lines in
\cref{fig:estwmig}), depending on how maladapted migrant individuals are.
This suggests that maladaptive gene flow in diploids can increase the
probability of tetraploid establishment.
However, at the same time we find that the tetraploid establishment probability
declines monotonically with $m$, which highlights the negative effects of MCE.
Together, these results confirm the view that, while cytotype differences are an
effective barrier to maladaptive gene flow, this advantage to establishment is
outweighed by the negative effects of MCE.

Our simulations further show that the mechanism of unreduced gamete formation
(as determined by the $\alpha_2$ parameter) can affect the establishment
probability (\cref{fig:estwmig}B, dashed lines).
This is mainly because the phenotypic variance of a newly formed tetraploid is
increased by a factor $(1+\alpha_2)$, thereby increasing the chance that a
newly formed tetraploid is well-adapted to the marginal habitat.
The rate of double reduction ($\alpha_4$) has a more limited effect
(\cref{s-fig:suppdr}).

Established diploid populations are more inbred on average than established
tetraploids when migration is weak, but the difference is slight except when
there is no migration at all (\cref{fig:estwmig2}, top row).
For stronger migration ($m > 0.1$), the opposite holds.
This is a result of two interacting processes.
On the one hand, inbreeding is slower in tetraploids, so that during adaptation
and establishment from a single or limited number of barely inbred individuals, 
the inbreeding coefficient is expected to increase less rapidly conditional on
the event of tetraploid establishment.
On the other hand, migration mostly introduces unrelated diploids, which cross
more readily with diploids than tetraploids, so that migration reduces the
average relatedness in established diploid populations much more than in
established tetraploid populations.

Conditional on establishment, tetraploids have a higher trait mean both in the
absence of migration and for moderate to strong migration, but not when
migration is weak (\cref{fig:estwmig2}, bottom row).
In the absence of migration, this is a consequence of the reduced rate of
inbreeding and the resulting increased adaptive potential of tetraploids.
Another consequence of this is that, in the absence of migration, the average
trait value in tetraploids at low population size (conditional on eventual
establishment) is lower than that in diploids.
This is because, in order to adapt before genetic variation is exhausted,
tetraploids need not increase the trait mean as much as diploids do in the
first couple of generation.

It is clear from \cref{fig:estwmig2} that migration is not only maladaptive,
but also introduces genetic variation upon which selection can act, thereby
increasing the mean trait value at the time of establishment for both
tetraploids and diploids.
Interestingly, when migration is strong, tetraploids have markedly larger
phenotypic values than diploids ($m=1$ in \cref{fig:estwmig2}).
Diploids suffer strongly from maladaptive gene flow when the population size is
low, while tetraploids enjoy some reproductive isolation.
Furthermore, in these replicates, tetraploids tend to emerge and rise in
frequency at larger population sizes on the island, and hence tend to derive
from diploids that already experienced several generations of selection.
Neotetraploids deriving from parents with $z>0$ will have more extreme
phenotypes on average (see methods) and hence be better adapted.

\subsection*{Loss of self-incompatibility and assortative mating}

We now briefly examine the effects of prezygotic isolation mechanisms that may
be associated with polyploidization and that could alleviate MCE.
Plausible sources of prezygotic isolation include assortative mating by
cytotype and self-fertilization.

When polyploidization disrupts an existing SI system (see e.g.
\cite{robertson2011comparative,zenil2019,novikova2023}), we expect
that tetraploids suffer less from MCE, as an appreciable portion of their
ovules are now assured to be fertilized by diploid gametes, irrespective of the
composition of the population.
At the same time, we expect that accelerated inbreeding in selfing tetraploids
diminishes the adaptive advantage of tetraploids that we found above.
As shown in \cref{fig:selfing}, we find that when polyploidization is
associated with the loss of a SI system (i.e. when diploids are
self-incompatible, but tetraploids are not), tetraploids have a strongly
increases establishment probability.
This is the case even when the selfing rate $\sigma$ in tetraploids is zero (in
which case, under our modeling assumptions, there is only random selfing, i.e.
the \textit{realized} selfing rate in tetraploids is $1/N$).

Besides reducing MCE, selfing is also expected to simultaneously reduce the
effects of maladaptive gene flow, so we would expect that tetraploid
establishment is promoted with increasing migration rates. 
However, increased migration would at the same time alleviate the issue of
reproductive assurance in diploids, as a diploid migrant is more likely to find
a compatible mate.
We find that when the selfing rate is sufficiently high ($\ge 0.4$ in
\cref{fig:selfing}A), the relative establishment probability of tetraploids
increases with increasing migration rate.
In this regime, the increased strength of MCE due to increased migration and
increased reproductive assurance in diploids is compensated by the stronger
maladaptive gene flow experienced by diploids.

Self-incompatibility is clearly a strong disadvantage when colonizing a novel
habitat, as a self-incompatible population of size one can never reproduce.
However, even when diploids are self-compatible, polyploids may still have
increased rates of self-fertilization (for instance due to altered flower
morphology). As an illustration, \cref{fig:selfing}A also shows results where
diploids are assumed to be self-compatible with $\sigma=0$ in diploids (i.e.
random selfing, dashed transparent lines).
The tetraploid establishment probability is still markedly increased when
$\sigma \ge 0.4$, and as for the simulations with self-incompatibility,
migration still promotes the probability of tetraploid establishment when the
selfing rate in polyploids is sufficiently large.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{/home/arzwa/dev/InfGenetics/doc/img/prez.pdf} 
    \caption{
    (A) Establishment with recurrent migration selfing in polyploids. 
    The solid lines show the case where diploids are self-incompatible 
    The dashed transparent lines show the case where diploids do random
    self-fertilization (i.e. self-fertilization occurs with probability $1/N$),
    Triploids and tetraploids have the same selfing rate. 
    $\sigma=0.0$ refers to random self-fertilization.  
    (B) Establishment with recurrent migration and assortative mating by
    cytotype. The rate of assortative mating is determined by $\rho_k = \rho$
    for $k=2,3,4$, where $\rho_k$ is the probability that an ovule from a
    $k$-ploid mother is pollinated by a $k$-ploid father.
    All results are based on 25.000 replicate simulations.  We assume
    $\gamma=0.25, \theta=1.5$ and $u=v=0.05$.  \label{fig:selfing}}
\end{figure}

Another prezygotic isolating mechanism that has often been considered relevant
for explaining tetraploid establishment is assortative mating by ploidy level,
where ovules from a tetraploid are more likely to be fertilized by
pollen coming from a tetraploid -- irrespective of the trait values of these
individuals.
Clearly, assortative mating increases the probability of tetraploid
establishment (\cref{fig:selfing}B), although not as strongly as the loss of an SI
system does.
Again, we find that for some parameter values (roughly $\rho \ge 0.4$),
assortative mating may be strong enough so that tetraploid establishment
increases with increasing migration rates, suggesting that tetraploids evade
maladapive gene flow sufficiently to overcome MCE.

\section*{Discussion}

The observation that polyploid populations tend to inhabit more extreme
habitats or occur at the edge of the range of their conspecific diploids
has spurred considerable interest among botanists and evolutionary biologists
\citep{rice2015,kolar2017,vandepeer2021,griswold2021,mortier2024}.
An important question is whether such patterns emerge because polyploids are
somehow more tolerant to extreme environmental conditions (i.e. they somehow
are intrinsically more fit than diploids at the range margin), or whether other
aspects of the population dynamics of mixed-ploidy populations may favor the
establishment of polyploid subpopulations in peripheral habitats.

In this study, we worked out the infinitesimal model for an additive polygenic
trait in autotetraploids and mixed-ploidy populations.
We used the mixed-ploidy infinitesimal model to study the establishment of
tetraploids in a marginal habitat by means of individual-based simulations. 
Assuming the trait to be under directional selection in the marginal habitat,
and migration of maladapted individuals from a predominantly diploid source,
we sought to determine the conditions under which tetraploids are more likely
to establish a stable population.
Importantly, we assume no intrinsic advantage of polyploids in the marginal
habitat, i.e. the expected fitness of a migrant individual is the same
regardless of the ploidy level.
Differences in establishment probabilities are hence caused solely by aspects
of autopolyploid genetics and the barrier to gene flow between subpopulations
of different ploidy levels.

When migration is weak, succesful establishment is not affected by
maladaptive gene flow and we can treat establishment in the marginal habitat as
independent trials of founding a population from a single individual.
In order to avoid extinction, the population has to increase the trait mean
by a sufficient amount before the genetic variation carried by the initial
migrant individual is exhausted.
The probability that the population manages to do so depends on the degree of
maladaptation, the intensity of selection and the rate of inbreeding.
We find that the decreased rate of inbreeding in autotetraploids gives a
rare tetraploid migrant a larger adaptive potential than a diploid migrant,
even if the genetic variance carried by the founding individual is the same.

In the presence of maladaptive gene flow, a nascent tetraploid subpopulation
suffers from MCE, and although polyploids are more reproductively isolated from
a typical migrant (and hence suffer less maladaptive gene flow), MCE will
preclude polyploid establishment as the rate of migration grows.
Nevertheless, it is important to remark that, the probability of tetraploid
establishment in the marginal habitat can be an order of magnitude higher than
expected based on the frequency of tetraploid migrants (i.e. is roughly $O(u)$
instead of $O(u^2)$) when migration is sufficiently weak and maladaptation
sufficiently high.

Additional sources of prezygotic isolation such as selfing and assortative
mating by cytotype may further boost the probability of tetraploid
establishment.
These processes interact with the rate of migration, so that when
selfing/assortative mating occurs above some threshold rate, the tetraploid
establishment probability increases with increasing migration rates, whereas
below the threshold it decreases with increasing migration pressure.
In the latter case, the advantage that tetraploids have when it comes to
avoiding maladaptive gene flow is not strong enough to overcome the effects of
MCE, whereas in the former case it is.

A major weakness of the present work is that we have ignored inbreeding
depression and dominance throughout.
Including dominance in the infinitesimal framework is already challenging for
diploids (requiring the tracking of 4-way identity coefficients;
\cite{barton2023}), and appears intractable for higher ploidy levels.
Nevertheless, autopolyploidy has important consequences whenever dominance is
relevant, as in the case of inbreeding depression.
Indeed, when inbreeding depression is due to recessive deleterious variation,
it should be less expressed in polyploids simply because, for the same allele
frequency, homozygous genotypes are much rarer in polyploids than in diploids
at Hardy-Weinberg equilibrium.
When there is indeed deleterious recessive variation, inbreeding during the
establishment process should therefore incur a higher fitness cost in diploids
relative to tetraploids, and hence further increase the probability of
tetraploid establishment.
In addition, inbreeding itself occurs at a slower rate in polyploids, further
decreasing inbreeding depression during establishment.

Dominance and inbreeding depression may strongly affect the complicated
relationship between selfing, migration load and tetraploid establishment
probability.
\cite{griswold2021} performed a study somewhat similar to ours, but did include
a model of inbreeding depression.
The author focused on the case where local fitness is determined by a single
biallelic locus, and investigated the interaction between inbreeding depression
and migration load (where inbreeding depression is modeled as an immediate
fitness reduction in offspring produced by selfing).
In his model, inbreeding depression is different between cytotypes, so that
tetraploids are able to produce more offspring through selfing relative to
diploids, who have to rely more on outcrossing.
However, outcrossing incurs maladaptive gene flow, and thereby puts the
diploids at a disadvantage. 
He found that autotetraploids can establish when adaptation in the
peripheral habitat is conferred by recessive alleles (so that migration load is
expressed when migrant alleles are rare) and when inbreeding depression in
tetraploids is lower than in diploids.
It would be very interesting to combine the infinitesimal framework with a
simple model of inbreeding depression to investigate in a somewhat more
realistic model whether the combination of maladaptive migration and
differential inbreeding depression could explain the prevalence of polyploid
subpopulations at range edges.

In the long term, polyploids are expected to accumulate a larger
mutation load when deleterious variation is recessive due to less efficient
purging, and this may yield \textit{increased} inbreeding depression.
These effects have been studied in the context of range expansions
\citep{booker2024}.
However, this applies only to polyploids that have been established for a long
time.
In our case, polyploids are always recently descended from diploid ancestors,
so they will not have accumulated more deleterious recessives than their
diploid counterparts, in which case polyploidy would lead to reduced rather
than increased inbreeding depression (as discussed above).

It would be very interesting to experimentally assess whether typical
quantitative traits in mixed-ploidy populations follow approximately the
mixed-ploidy infinitesimal model we outlined in this paper, and to estimate the
relevant model parameters (in particular the $\beta_3$ and $\beta_4$ values).

\section*{Acknowledgements}

I thank Christelle Fraïsse for employment, encouragement and feedback. I
acknowledge funding from the Research Foundation -- Flanders (FWO, Junior
Postdoctoral Fellowship 1272625N).

\bibliographystyle{abbrvnat}
\bibliography{/home/arzwa/vimwiki/bib.bib}
\end{document}