### 2025-03-13 doc/paper/titlepage.tex
starting revisions

\documentclass[12pt,a4paper]{article}
\input{preamble.tex}
\usepackage[margin=1in]{geometry}
\usepackage{authblk}
\externaldocument[s-]{supp}

\begin{document}
\title{Autopolyploid establishment through polygenic adaptation}
\author[1]{Arthur Zwaenepoel\thanks{arthur.zwaenepoel@uantwerpen.be}}
\affil[1]{\footnotesize Department of Biology, University of Antwerp, 2020 Antwerp, Belgium}
\date{\vspace{-5ex}}
\maketitle

\subsection*{Data availability}

Individual-based simulations for the mixed-ploidy infinitesimal model were
implemented in Julia \citep{julia}.
Documented code and simulation notebooks are available at 
\url{https://github.com/arzwa/InfGenetics}.

\subsection*{Funding}

I acknowledge funding from the Research Foundation -- Flanders (FWO, Junior
Postdoctoral Fellowship 1272625N).

\subsection*{Conflict of interest}

I have no conflict of interest to declare.

\subsection*{Acknowledgements}

I thank Christelle Fraïsse for encouragement and feedback. Thanks to Quinten
Bafort, Felipe Kauai and Frederik Mortier for comments on an early draft
version.

\bibliographystyle{abbrvnat}
\bibliography{/home/arzwa/vimwiki/bib.bib}

\end{document}