### 2025-04-03 doc/paper/response.md
update

## Associate Editor

>Thank you for submitting your article to Evolution. It has now received two
reviews, both of which recognize its valuable contributions and essentially
recommend publication after relatively minor revision. I share these views.
While both reviews are positive, they make a number of worthwhile suggestions
for you to consider for revising the manuscript. One reviewer has suggested
combining the Results and the Discussion into a single section to avoid
unnecessary discussion of the results in the current Results section.
Personally, I would favor retaining both sections, **moving more discussive text
from the Results to the Discussion**. I leave the decision to you, but I do agree
with the reviewers that some resorting of the Results text is needed. I also
agree that the **Introduction could be better motivated by referring more to
empirical work**, and the **Discussion section could relate the results more to
other theoretical findings** in the literature. Overall, think that revision will
be relatively straightforward, but I would ask that you consider all the
suggestions made by the reviewers and, if you disagree with them, that you
justify and explain your reasons in a response letter. (emphasis mine)

Thanks a lot for your positive appraisal.

First of all, I wish to note that there was an important error in my original
submission, which I outline below. Evidently, I have redone all simulations and
revised the manuscript accordingly. The error affects both diploid and
tetraploid establishment similarly, so does not affect the results
qualitatively (since these are concerned, mostly, with relative establishment
probabilities), but does affect all results with nonzero selection intensity
($\gamma > 0$). I outline the details of the error in the next section.

I have now revised the Introduction to refer more to empirical work.
I have included the references suggested by Reviewer 1.
Following the suggestions of Reviewer 2, I have tried to refer better to
empirical work on (1) polyploid establishment at range margins and (2) the
genetics of adaptation to peripheral environments (in polyploids).
In the interest of brevity, I do not present an in-depth review of these
topics. For (1), I refer mostly to @griswold2021, who presented a
nice overview. For (2), I mainly tried to better motivate my interest in
*polygenic* adaptation, referring to recent work and reviews that suggest that
local adaptation is typically polygenic, and work that shows that
autopolyploids present no exception to this.

Regarding the Results and the Discussion sections: I do not want to do away
with Discussion section altogether (as suggested by Reviewer 1), as I
wish to discuss factors relevant to the problem which are not studied in my
article (dominance, inbreeding depression, *etc.*) and hence do not feature in
the results section. Furthermore, I want to relate my work to other recent
theoretical work (in particular that of @griswold2021). I believe this warrants
a dedicated Didcussion section.
I have sought to move 'more discussive text' to the Discussion section,
although I found there was fairly limited scope to do so when I seek to keep
the interpretation of the results more or less self-contained.

I have expanded the discussion slightly to address some of the remarks of
Reviewer 2 concerning the issues of mutation and the scaling of genetic variance. 
I have added a concluding paragraph following the suggestion of Reviewer 1 to
highlight that other eco-evolutionary questions about mixed-ploidy populations
might be fruitfully addresed using the infinitesimal framework. I took the
opportunity to allude to the work of @oswald2011 in that section, following up
on a suggestion of Reviewer 2.

Overall, I think these revisions have contributed considerably to a
better-motivated article that engages more with relevant empirical work. I
hence thank the reviewers and the editors for their suggestions. Below I
include a detailed reply to both reviewers.

## Error in the first submission

In the simulations included in the first submission of my article, I did not
sample the trait values of the next generation correctly when there was nonzero
selection $\gamma > 0$ (as in figs. 2-5).
As outlined in my paper, I sample the number of offspring of a particular
combination of gametes from a particular parental pair after selection in
proportion with the expected fitness using eq. 11 (number in new version).
In the previous version, I then sampled, incorrectly, offspring trait values
from the Gaussian distribution of offspring implied by the mixed-ploidy
infinitesimal model (eq. 7). This however fails to account for selection on the
trait value within each parental pair.

The density of trait values after selection is found easily.
Writing $f(z)$ for the Gaussian density of trait values with mean $\bar{z}$ and
variance $V$, and $w(z)$ for the fitness function, the density after selection
$f^\ast(z)$ is
\begin{align*}
f^\ast(z) &= \frac{w(z)f(z)}{\int w(z)f(z)dz} \\
&\propto \exp\left(\gamma(z - \theta) - \frac{(z - \bar{z})^2}{2V}\right) \\
&\propto \exp\left(-\frac{1}{2V}(z^2 - 2z\bar{z} - 2V\gamma z)\right) \\
&= \exp\left(-\frac{1}{2V}\left[z^2 - 2z(\bar{z} + \gamma V) + (\bar{z} + \gamma V)^2 - (\bar{z} + \gamma V)^2\right]\right) \\
&\propto \exp\left(-\frac{1}{2V}\left[z - (\bar{z} + \gamma V)\right]^2\right)
\end{align*}
i.e. $f^\ast(z)$ is a Gaussian density with mean $\bar{z} + \gamma V$ and
variance $V$. 
The correct trait value distribution for the offspring in parental pair $(i,j)$
with gametes of ploidy $k$ and $l$ *after selection* is hence a Gaussian with
mean $\overline{z_{ij}^{kl}} + \gamma V_{ij}^{kl}$ and variance $V_{ij}^{kl}$.

The incorrect results from the previous submission can be seen as a kind of
zeroth order approximation to the correct results: i.e. they implement
selection on the mean trait value across families, but not selection on the
variance. The main consequence of this error is that adaptation is more
efficient than in my previous results, and hence establishment is easier. This
is however the case for both diploids and tetraploids. 
The effect on establishment is however nonlinear, so although the main qualitative
results of the paper remain the same, the new results do not simply amount to a
rescaling of the old ones.

I apologize for this oversight.

## Reviewer 1

>This interesting manuscript investigated how the infinitesimal model can be
expanded to polyploidy populations (triploids and tetraploids here). It
proposes a mathematical formulation of a polyploid population's expected
genetic diversity for a given amount of inbreeding. The model is then used to
study how polyploidy can affect the adaptation to a new habitat (based on a
sink-source model).
I found the manuscript to be interesting and only have a few things to say
about it. I will for sure motivate further expansion of the model to understand
the evolution of autopolyploidy.
Having no line number does not help to comment precisely on the manuscript; I
did my best to be as precise as possible.

Thank you for these encouraging words, and my apologies for the oversight
when it comes to line numbers!

>Results are already well discussed when presented. I would either merge the
results and discussion sections (as in brief communications for example),
because it reads well as presented, or remove any explanations/discussion from
the results to put them in the discussion section. The first option sounds
better to me.

I appreciate the comment and have tried to rework the text with this in mind.
However, following the suggestion of the associate editor, I went with the
second option. I refer to my answer to the associated editor above.

>A small concluding section where the authors mentioned other questions to
study with their model will be useful in my opinion. The author proposed to
complexify the current model of adaptation with dominance or to relax some
hypotheses, but I think other biological scenarios could be studied too.

Whereas I don't think listing potential biological scenario's that can be
investigated using my modeling approach is very useful, I think the reviewer is
right in that it may be useful to highlight that the modeling approach can be
employed to address other questions related to the ecology and evolution of
mixed-ploidy populations. I have now added a small paragraph at the end that
does so.

>Defining establishment when $N=100$: Is it a classic way to define
establishment in such models?  Or does it come from your experience with the
model and you see that Under N=100 extinction remains likely? More details
would be needed here.

$N=100$ is indeed a fairly arbitrary choice. This threshold was also adopted in
@barton2018, so it makes comparisons of establishment probabilities *etc.* with
their paper somewhat more straightforward.
$N=100$ seems to be a reasonable threshold in that extinction becomes
sufficiently unlikely, while we do not have to simulate large populations (our
approach requires tracking the $N \times N$ matrix of identity coefficients). 
Also, for the migration rates considered, $N=100$ is sufficiently large that
migration alone (without population growth due to adaptation) can not lead to
$N=100$, i.e. the population does have to increase the mean trait value in
order to reach establishment.
I did a check by putting the threshold at $N=150$, and the establishment
probabilities do not change appreciably.

>Abstract: I am not sure the reference to Barton et al. (2017) is appropriate
here. Maybe a more general definition would fit better.

The reason for the reference is that there appears to be some confusion about
different models that bear the name 'infinitesimal model' (as outlined in for
instance @turelli2017 and @walsh2018). I wanted to make it clear that we use
the "Gaussian descendants" infinitesimal model, which is very carefully
outlined in @barton2017. However, I guess this is sufficiently clear from the
introduction and the methods sections, so the reference in the abstract can
indeed be omitted, as it is in the revised manuscript.

>Introduction: P2 (lower inbreeding depression (ronfort 1999, Otto & Whitton
2000): Some empirical data?  Husband et al. (2008), Clo & Kolar (2022).

I have now added the suggested references, thank you.

>Model and methods: P3-P4 (mixed ploidy population model): maybe write that the
rate of reduced gametes is fixed and genetically inherited from diploids to
tetraploids. It is clear from the equations but I suspect it could be clearer
for non-theoretical people who will not necessarily focus that much on
equations or matrices.

I added a sentence below eq. 1 to emphasize this. I note that this is also
emphasized in the discussion "In addition, we have assumed a relatively high
and constant rate of unreduced gamete formation $u$ and triploid fertility $v$
in all our simulations (5%), whereas these are known to be variable across the
population, and at least in part genetically determined (Kreiner et al., 2017a;
Clo et al., 2022)"

>P4 (tetraploids are not twice as big as diploids): Some references would be
needed to justify (Porturas et al. 2019 for example).

I have added the reference to Porturas et al. (2019).

>Discussion about inbreeding depression: again, some empirical data are
available to support some statements (Husband et al. (2008), Clo & Kolar
(2022)).

I have added the suggested references.

## Reviewer 2

>This paper is an important foundational work in the area of theoretical polyploid
evolutionary genetics. It presents an infinitesimal model of trait evolution for
triploids and tetraploids, and then applies the model to assess conditions under
which a tetraploid population will establish in a peripheral region connected to
a central core by dispersal. In terms of the infinitesimal model, several aspects
are very helpful

>
- Demonstrating how genetic effects are scaled across triploids and tetraploids
such that segregating variance is equivalent across ploidies, including with diploids
- How to combine the genetic effects in an individual with mixed-ploidy ancestry
- Calculating inbreeding and co-ancestry values for mixed-ploidy individuals
- Calculating the effect of unreduced gamete formation on segregation variance
across diploids, triploids and tetraploids

>Other informative analyses include recursive or Markov models of null
expectations for frequencies of ploidies, times to diploid ancestry and
effective population size across a population of mixed-ploidy.
Besides this set of results, the paper then examines conditions under which a
tetraploid population is established in a peripheral population connected to a
central core population that is predominantly composed of diploids. Given the
additive nature of gene-action in the infinitesimal model, this analysis will likely
be of interest to empiricists because additive effects are thought to be the least
constrained basis for something to evolve and there is empirical evidence that
the genetic basis of polyploid establishment in some species is additive. This is
in contrast to the theoretical results of Griswold (2021), which highlighted the
potential for a recessive-basis for autotetraploid establishment in a peripheral
region.
Main comments focus on mutation and adding background/context.

I thank the reviewer for this appraisal.

>Mutation:
For low migration rates, the possibility that autotetraploid relative to diploid es-
tablishment is driven by differences in how inbreeding accumulates is interesting
(figure 4), but the model does not include mutation. New mutation dissociates
identity by descent from identity in state and likely slows the rate of loss of
segregation variance. It would be helpful to address how scaled mutation may
affect dynamics of segregating variance between diploids and autotetraploids
and whether mutation may affect results. Is there a relationship between $\mu V_m$
(Barton et al. 2017) and $m$ and $V$ on establishment?
Recognizing the point seems to be important, but its analysis can be for later.

I have not considered mutation at all so far because I assumed it would not be
relevant for the population sizes and timescales considered in my study.
Specifically, any individual at the time of establishment derives from a
completely outbred migrant individual a relatively short time in the past (10
to 100 generations, say), so the contribution of new mutation to differences in
establishment probability between diploids and tetraploids should be
negligible. In more quantitative terms: we assume $1/\mu \gg \tilde{T}$ where
$\tilde{T}$ is the time between establishment and the arrival of the migrant
from which the established population derived. 
In the presence of migration the variance contributed by mutation is even more
certainly negligible.
As long as $m V$ is sufficiently larger than $\mu V_m$, we will have that, for
every mutation contributing a very slight decrease in the loss of segregation
variance on the island, there will be many more migrant arrivals, which each
introduce a completely unrelated genotype.
In all our simulations with migration we assume $mV$ considerably larger than
conceivable values of $\mu V_m$.

Of course, when considering the mixed-ploidy infinitesimal model independently
of the establishment question, mutation does become a relevant topic. I mention
this in the revised discussion but I think it is out of scope to discuss this
in any detail in the paper. I think that my approach for scaling across ploidy
levels that is explicitly based on underlying allelic effects in a finite-locus
model extends readily to mutation. The main difference between ploidy levels is
of course that the mutational target size is proportional to the ploidy level,
so that, unless the mutational variance is exactly halved in tetraploids, the
contribution of mutation to genetic variance ($\mu V_m$) in tetraploids is
larger than in diploids.

I have added a note in the discussion regarding mutation and why I have ignored it.

>In the intro providing empirical examples of polyploid establishment in
peripheral habitats would broaden readership and connections to the
literature. Also, reviewing empirical work on the genetics of adaptation to
peripheral environments in polyploids would broaden connections to the
literature.

I agree with the relevance of empirical work on polyploid establishment and
adaptation to marginal habitats for the introduction of the paper. However, I
think that to present an actual review of empirical studies here is a bit out
of scope. Griswold presents a brief review in his 2021 paper, and I prefer to
refer to his paper instead of paraphrasing his paragraph. Specifically, I now
write:
"Many empirical studies of mixed-ploidy populations find that polyploids
established in peripheral habitats at the edge of a species’ range (reviewed in
Griswold (2021)), and this is in accord with large scale biogeographical
patterns (Rice et al., 2019)".

While I think it is out of scope to review in any depth the empirical
literature on adaptation and polyploidy, I agree that the paper lacked a clear
connection to such work, and in particular I noticed that it lacked a clear
motivation for the interest in *polygenic* adaptation.
I have now tried to better motivate the likely polygenic nature of local
adaptation and have added a couple of references to recent work on the genetics
of adaptation in autopolyploids: 
"More often than not, local adaptation is polygenic in nature (Pritchard and Di
Rienzo, 2010; Barghi et al., 2020; Bomblies and Peichel, 2022), involving many
weakly selected variants across the genome, and adaptation during polyploid
establishment is unlikely to be an exception. Recent studies on local
adaptation in autopolyploids indeed tend to find a polygenic basis of
adaptation (Bohutı́nská et al., 2021; Konečná et al., 2021, 2022), however it is
not clear how observed adaptive differentiation in established tetraploid
populations relates to adaptation that occurred during initial establishment".

>The discussion could relate findings to other theoretical works more. For
example, it does not go back and place the results in the context of Oswald and
Nuismer (2011), which assumed a two-locus model and varied dosage effects, as
well as competition between diploids and tetraploids, including no competition.
In addition it allowed for assortative mating.

I agree that @oswald2011 could be a relevant paper to relate my work to,
however it should be noted that this paper focuses on establishment in
sympatry, and does not deal with migration, nor establishment in a new
environment (they do study establishment after environmental change, but as far
as I can tell they assumed a constant population size, so this is quite
different from establishment in a new habitat).
While my study is clearly related to theirs in that it is concerned with
establishment of autotetraploids in mixed-ploidy populations, it is not
straightforward to explicitly relate the results beyond noting that (1) they
investigated more processes (assortative mating by trait value, competition,
stabilizing selection and inbreeding depression) and (2) also found that
assortative mating and selfing may promote establishment.  The latter was of
course already found by many authors before @oswald2011...

However, I have now added a more explicit reference to @oswald2011 in the concluding
note (see the associated remark of Reviewer 1), to highlight that it would be
possible to do a study that is comparable to theirs but assuming infinitesimal
quantitative genetics instead of a model with a small number of loci.

>It would be informative to expand on the the consequences of different
scalings of genetic effects. This paper made the sensible choice to scale such
that segregation variance is equal between ploidies as a baseline. Griswold
(2021) scaled genetic effects such that autotetraploid and diploid individuals
have equal fitness as a function of proportional allele count in an individual.
A consequence of Griswold’s scaling is autotetraploids have lower additive
genetic variance at HWE (Gallais 2003, p. 185), so are expected to respond more
slowly in the absence of another process. Gallais (2003, p. 186) mentions some
studies that compared additive genetic variance between diploids and
autotetraploids.

The consequences of different scalings is taken up in the results and
discussion throughout the paper where I believe this is relevant.
In the methods section the relationship between allelic effects and additive
genetic variance is stated explicitly.
In the results shown in Fig. 2, the effect of scaling assumptions on
establishment probability is explicitly shown and described.
In the section on establishment with recurrent migration, the phenomenon that
polyploid offspring is more extreme on average when allelic effects are not
exactly scaled by one half (as in Griswold) is taken up. 
I now added some further discussion, referring explicitly to Griswold's scaling
in the Discussion section. I now also refer to @porturas2019 and @gallais2003.

>Neither the intro nor the discussion revisit what seems to be an important point
in Barton & Etheridge (2018, p. 111) about local adaptation via infinitely small
effects.

I assume the reviewer is referring to the following issue: classical single
locus population genetics predicts that one needs $m < s$ for selection to
maintain a locally beneficial allele when migrants introduce the deleterious
allele. So, when adaptation is due to alleles of very small effect, this seems
to imply that there is little scope for local adaptation, since locally
beneficail variants will be swamped by gene flow.

There are multiple reasons why this is not the case in nature. The argument
outlined in @barton2018 is simply that the classical theory usually assumes
complete divergence between mainland and island, but that when adaptation is
from standing variation instead, this does not hold, since allele frequency
differences between mainland and island are only slight. In the latter case,
many alleles of small effect can be divergently maintained by selection.
Another reason why this does not hold is that selected alleles will 
be in linkage disequilibrium, so that entire sets of deleterious alleles
introduced by migrants are eliminated jointly. The latter phenomenon is
explored in detail in @sachdeva2022 and @zwaenepoel2024.

I do not think, however, that it is pertinent to discuss these issues in the
present paper -- they are somewhat technical and require quite some space to
explain and do not have much to do with the problem studied, i.e. polyploid
establishment, so in the interest of brevity I decided not to revisit this
point.

### Other comments:

>p.1, Abstract: Use the word “found”, but is “establish” more accurate? Overall,
the corresponding sentence is difficult to understand/parse. Suggest revision.

This sentence has been revised.

>p. 1: “negative frequency dependence” -> Isn’t MCE a positive-frequency de-
pendent process?

Indeed, it is! This has been corrected.

>p. 2: Suggest citing Bulmer (1971).

I am unsure whether this is a good idea, as it may lead to some confusion as to
which version of the infinitesimal model we are working with.
The model described in Bulmer (1971) is not exactly equivalent to the Gaussian
descendants approximation that I use (as outlined in Turelli (2017), which I
cite).
If the reference is suggested by the reviewer in order to cite a more
'original' reference, I don't think there is a reason to single out Bulmer
(1971), one could refer to Fisher (1918) or work by Alan Robertson in the
1960's.
I think the reference to Turelli (2017), which reviews the history of the
infinitesimal model, is quite adequate, and prevents confusion as to what we
mean when we refer to 'the infinitesimal model'.

>p.7: Suggest adding “, across gene copy arrangements and ploidies” at the end
of the last sentence before the “Establishment” section.

I do not fully understand this suggestion. I personally find this makes the
sentence more confusing, as I never talk of gene copy arrangements.

>p. 7, Establishment section: The organization of paragraphs two and three in
this section is a bit confusing. N is used before it is formally defined. In my
first reading I thought there was migration and then selection, whereas there is
selection and then migration. It would also help support a self-contained paper
to give the expression for $E\left[e^{\gamma(\dots)}\right]$.

As far as I can tell, $N$ is defined when first mentioned. I slightly altered
the sentence to make this more clear.

I added the expression for the expected fitness (eq. 11).

The life cycle is indeed as follows: migrants arrive, they mate randomly and
offspring survives with a probability that is determined by its fitness. This
is indeed 'selection before migration'. I tried to make this more clear.

>p. 11, figure 3 (and later figure 5): The smallest value of m is 0.1 in these
figures, but based on figure 4, it looks like m = 0.01 is a transition point from
processes shaped by the accumulation of inbreeding versus migration becoming
more of a factor. Can m be extended to 0.01?

I extended the relevant figures as suggested, as $m=0.01$ is perhaps more
representative of weak migration than $m=0.1$
Weak migration means essentially that $m^2$ is negligible, so migrants and
early generation descendants thereof do not mate with migrants and the whole
process amounts more or less to independent establishment trials (there is no
swamping).

I do not think, however, that fig 4 suggests such a transition point though. As
noted in the main text the difference between the $m=0$ and $m=0.01$ results in
fig. 4 are mostly due to the fact that the latter assumes a Gaussian
distribution of migrant trait values, whereas the former assumes a fixed
initial migrant trait value.
(*"The similar trait means for weak migration suggest
that establishment depends mostly on the chance pick of a well-adapted migrant
(note that the $m = 0$ simulations in fig. 4 assume the initial migrant has $z =
0$, whereas the $m = 0.01$ simulations assume the initial trait value to be
Gaussian)."*)

>p. 12: Add qualifier “as $m$ increases” at end of last sentence in last full
paragraph.

Done.

>p. 14, Figure 5 caption: “migration and selfing”

Fixed.

>p. 4, suppl., figure S4: “formation”

Fixed.

>p. 6, suppl., equation 6: Should the square only be in the denominator?

That's right, thank you!

>p. 10, suppl.: “coancestry coefficients in (10) is “ - missing the “(10)”

Fixed, thank you!

>p. 12, suppl.: add “between centromere and locus” after “... recombination
happens with probability $c$”

Done.

>p. 13-15, suppl.: Some notation seemed to not be defined, e.g. $V_{S(2,2)}$, ...

The notation is introduced on p.11 in the beginning of S2.6. I have slightly
reorganized the first paragraphs of S2.6 to make the notation and general
approach more clear.

## References